### 2017-06-17 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Creating R Markdown
This will help us to aggregate everything into one place. Prof Holt mentioned this in class.

---
title: "MSDS 7333 Case Study 06"
author: "Matthew Baldree, Ben Brock, Tom Elkins, Austin Kelly"
date: "June 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assignment: 
1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both: 
    + chosenAP = "00:0f:a3:39:e1:c0"
    + rejectedAP = "00:0f:a3:39:dd:cd"
* *Which of these two MAC addresses should be used and which should not be used for Real Time Location Services (RTLS)?*
* *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
* *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*

***

2. Implement alternative k-nearest precition method using weights on received signal strength.
* *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
* *Use calcError() to compare approaches*

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-06-17 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
updates

@@
     + rejectedAP = "00:0f:a3:39:dd:cd"
-* *Which of these two MAC addresses should be used and which should not be used for Real Time Location Services (RTLS)?*
+* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
 * *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+## Introduction
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+***
 
-```{r cars}
-summary(cars)
-```
+## The Approach (Explained)
 
-## Including Plots
+***
 
-You can also embed plots, for example:
+### Preliminary Functions to be used:
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
+## The Output Achieved (this should be included as close as possible to the output or figures as possible)
 Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-06-18 MSDS7333-baldree-case6-fx.r
almost there for part 1

@@ -179,14 +179,188 @@ castOnline = function(df) {
   #   df: online data frame
   #
   keepVars = c("posXY", "posX","posY", "orientation", "angle")
+  numOfAPs = length(unique(df$mac))
   byLoc = with(df,
                by(df, list(posXY),
                   function(x) {
                     ans = x[1, keepVars]
                     avgSS = tapply(x$signal, x$mac, mean)
-                    y = matrix(avgSS, nrow = 1, ncol = 7,
+                    y = matrix(avgSS, nrow = 1, ncol = numOfAPs,
                                dimnames = list(ans$posXY, names(avgSS)))
                     cbind(ans, y)
                   }))
   return(do.call("rbind", byLoc))
 }
+
+createOfflineSummary = function(df) {
+  # In order to examine distribution for all locations, angles, and are two interested APs, we
+  # will create a summary statistics for all location-orientation-AP combinations with a new factor.
+  # For each combination there are around 100 observations.
+  #
+  # Args:
+  #   df: offline data frame
+  #
+  df$posXY = paste(df$posX, df$posY, sep="-")
+
+  # create data frames for each combination
+  byLocAngleAP = with(df, by(df, list(posXY, angle, mac), function (x) x))
+
+  # summary statistic
+  signalSummary = lapply(byLocAngleAP, function(oneLoc) {
+    ans = oneLoc[1, ]
+    ans$medSignal = median(oneLoc$signal)
+    ans$avgSignal = mean(oneLoc$signal)
+    ans$num = length(oneLoc$signal)
+    ans$sdSignal = sd(oneLoc$signal)
+    ans$iqrSignal = IQR(oneLoc$signal)
+    ans
+  })
+  return(do.call("rbind", signalSummary))
+}
+
+reshapeSS = function(df,
+                     varSignal = "signal",
+                     keepVars = c("posXY", "posX", "posY"),
+                     sampleAngle = FALSE,
+                     refs = seq(0, 315, by=45)) {
+  # reshape signal strength help function
+  # aggregate signal strengths from these angles and create a data structure similar to onlineSummary
+  # Args:
+  #   df: dataframe to reshape
+  #   varSignal: variable to aggregate
+  #   keepVars: variables to retain
+  #   sampleAngle: true if we want to select one angle at random from df
+  #   refs: angle references
+  #
+  numOfAPs = length(unique(df$mac))
+  byLocation = with(df, by(df, list(posXY),
[…]

### 2017-06-18 MSDS7333-baldree-case6.r

@@ -30,6 +30,9 @@ source("MSDS7333-baldree-case6-fx.r", print.eval=TRUE)
 # read offline data
 offline = readData()
 
+# create factor for all unique xy combinations
+offline$posXY = paste(offline$posX, offline$posY, sep = "-")
+
 # access points (AP) addresses in question
 # note that both devices are Alpha Networks devices because they begin with '00:0f'
 chosenAP = "00:0f:a3:39:e1:c0"
@@ -72,25 +75,8 @@ plotDensitySignalStrength(df)
 
 #### Average Signal Strength Distribution ####
 
-# In order to examine distribution for all locations, angles, and are two interested APs, we
-# will create a summary statistics for all location-orientation-AP combinations with a new factor.
-# For each combination there are around 100 observations.
-offline$posXY = paste(offline$posX, offline$posY, sep="-")
-
-# create data frames for each combination
-byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function (x) x))
-
-# summary statistic
-signalSummary = lapply(byLocAngleAP, function(oneLoc) {
-  ans = oneLoc[1, ]
-  ans$medSignal = median(oneLoc$signal)
-  ans$avgSignal = mean(oneLoc$signal)
-  ans$num = length(oneLoc$signal)
-  ans$sdSignal = sd(oneLoc$signal)
-  ans$iqrSignal = IQR(oneLoc$signal)
-  ans
-})
-offlineSummary = do.call("rbind", signalSummary)
+# create offline summary data frame
+offlineSummary = createOfflineSummary(offline)
 
 # plot stddev of average signal strength per targeted APs
 # result: reinforces that both addresses are important for full range coverage
@@ -107,7 +93,9 @@ plotStdDevSignalStrength(df)
 # points will be used to predict their coorindations and the error prediction will be determined
 # for each model.
 
+# look at online data
 macs = unique(offlineSummary$mac)
+macs
 online = readData("online.final.trace.txt", subMacs = macs)
 online$posXY = paste(online$posX, online$posY, sep="-")
 length(unique(online$posXY))
@@ -120,3 +108,45 @@ onlineSummary = castOnline(online)
 # confirm rejected AP is now included
 dim(onlineSummary)
 names(onlineSummary)
+
+# for k-NN, we will include training data with angles close to point in question since angle matters.
+# if we want one angle, then include angles that match the rounded orientation of new observation.
[…]

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd

@@
 date: "June 17, 2017"
-output: html_document
+output: html_notebook
 ---
@@
 
+### Part 1
+
+Before we begin, we will need to set the parameters for printing and formatting
+```{r}
+options(digits = 2)
+options(error = recover, warn = 2)
+```
+Before continuing, we will need to address another file which houses the predefined functions. This helps to save notebook real estate and prevents confusion:
+```{r}
+source("MSDS7333-baldree-case6-fx.r", print.eval = TRUE)
+```
+
+Read offline data
+```{r}
+offline = readData()
+```
+
+Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both:
+
+    * chosenAP = "00:0f:a3:39:e1:c0"
+    * rejectedAP = "00:0f:a3:39:dd:cd"
+
+Since we have the two MAC addresses in question, we will declare these addresses as variables to simplify later code.
+
+```{r}
+chosenAP = "00:0f:a3:39:e1:c0"
+rejectedAP = "00:0f:a3:39:dd:cd"
+remainingAP = unique(offline[!offline$mac %in% c(chosenAP, rejectedAP),]$mac)
+```
+
+* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
+
+#### Here we will create a Signal Collection Map
+For this exercise, we will plot both the count and locations of the signals recorded for the access points.
+Even though it would be ideal to plot the access points(APs) directly on the map, we do not know their X and Y coordinates.
+
+```{r}
+plotSignalMaps(c(chosenAP, rejectedAP))
+```
+
+At first glance, we can see that the maps are quite similar.
+Unfortunately, the information gained from a visual comparison of these two diagrams does not give us a significant insight into which MAC address should be kept and which should be rejected.
+The next step to reach our goal will be to look at the individual Signal Strength Distributions
+
+***
+
+#### Signal Strength Distributions
+Plotting the individual signal strength distributions for each AP angle will help add perspective to the information we are working with.
+For the sake of consistency, we will use the same stationary coordinates listen in the course book (2, 12) (Add reference here.)
+
+```{r}
+df = subset(offline, posX ==2 & posY == 12 & !(mac %in% remainingAP), c(signal, angle, mac))
+plotBoxplotSignalStrength(df)
+```
+
+From the literature, we know the higher the decibel-milliwatts means a stronger signal strength.
+<font color="red">**(consider revising this sentence.
+Seems incomplete.)**</font> A stronger signal strength indicates a closer position to the AP since signal strengths decrease exponentially with distance.
+Comparing the signal strength distributions in this chart, we see "e1:c0" has a stronger signal strength at this location than "dd:cd."
+
+What if we compared the two addresses from the opposite end of the building? 
+```{r}
+# Let's focus on the fixed point (33,3)
+df = subset(offline, posX == 33 & posY == 3 & !(mac %in% remainingAP), c(signal, angle, mac))
+plotBoxplotSignalStrength(df)
+```
+Through the use of another perspective, the MAC address with the weaker signal at (2,12) still has a relatively weak signal when viewed from an AP at the other side of the building.
+The evidence against the address with the lower signal strength is mounting.
+
+***
+
+#### Time to investigate Signal Density
+To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each angle.
+This will help display the defining characteristics for each address.
+
+```{r}
+df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
+plotDensitySignalStrength(df)
+```
+<font color = "red"> &uarr; Having some trouble displaying the individual graphs here. &uarr; </font>
+
+* *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
+* *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*
+
+### Part 2
+
 ***
@@
 ### Preliminary Functions to be used: 
+## Let's break this section up and explain each function individually
 

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Finished transcribing R-Markdown

@@
 
+#### Average Signal Strength Distributions
+To examine the distribution for all locations and angles of the two APs in question, we will generate the summary statistics.
+The summary statistics for all possible combinations of locations and orientations for each AP through the use of a new factor.
+
+```{r}
+offline$posXY = paste(offline$posX, offline$posY, sep="-")
+
+# create data frames for each combination
+byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
+```
+
+Now, we calculate the summary statistics:
+```{r}
+signalSummary = lapply(byLocAngleAP, function(oneLoc) {
+  ans = oneLoc[1, ]
+  ans$medSignal = median(oneLoc$signal)
+  ans$avgSignal = mean(oneLoc$signal)
+  ans$num = length(oneLoc$signal)
+  ans$sdSignal = sd(oneLoc$signal)
+  ans$iqrSignal = IQR(oneLoc$signal)
+  ans
+})
+offlineSummary = do.call("rbind", signalSummary)
+```
+
+The Standard Deviation of the average signal strength per targeted AP
+
+```{r}
+df = subset(offlineSummary, mac %in% c(chosenAP, rejectedAP), c(sdSignal, avgSignal, mac))
+plotStdDevSignalStrength(df)
+```
+
+#### Let's dive into the K-Nearest Neighbors:
+<font color = "red"><b>Discuss KNN here</b></font>
+
+```{r}
+macs = unique(offlineSummary$mac)
+online = readData("online.final.trace.txt", subMacs = macs)
+online$posXY = paste(online$posX, online$posY, sep = "-")
+length(unique(online$posXY))
+tabonlineXYA = table(online$posXY, online$angle)
+#output shows measurements were taken at all angles throughout the floor
+tabonlineXYA
+```
+
+Organize the data where each AP is in a column.
+
+```{r}
+onlineSummary = castOnline(online)
+#confirm rejected AP is included
+dim(onlineSummary)
+names(onlineSummary)
+```
+
 * *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added rest of Matt's work to markdown

@@
 
+<font color = "red">(Continued from baldree-case6.r)</font>
+
+> for k-NN, we will include training data with angles close to point in question since angle matters.
+> if we want one angle, then include angles that match the rounded orientation of new observation.
+> if we want two angles, then pick two multiples of 45 degrees that flank the new observation's orientation.
+> if we want three angles, then pick the closest 45 degree increment and on either side of it.
+> what if we didn't want to collapse the signal strengths across the m angles, and
+> instead return a set of mx166 signals for each access point?
+
+```{r}
+for (ap in c("", rejectedAP, chosenAP)){
+  test = onlineSummary[, !(names(onlineSummary) %in% ap)]
+  train = subset(offlineSummary, mac != ap)
+  err = predictionErrors(test, train, numberOfAnglers = 1, k = 3)
+  print(paste(ap, err))
+}
+```
+
+```{r}
+#Iterate through 1 to 3 neighbors with 1 to 3 angles for each scenario to determine best combination
+#for cross validation
+for (ap in c("None", rejectedAP, chosenAP)){
+  if (ap == "None"){
+    test = onlineSummary
+    train = offlineSummary
+  } else {
+    test = onlineSummary[, !(names(onlineSummary) %in% ap)]
+  }
+
+  estXY = predXY(newSignals = test[, 6:11],
+                 newAngles = test[, 4],
+                 train, 
+                 numAngles = 3, k = 3)
+  actXY = test[, c("posX", "posY")]
+  err = calcError(estXY, actXY)
+
+  print(paste(ap, err))
+}
+```
+
+<font color = "green" size = "5px"><b>This is where we will perform cross validation and plot for desired AP for a range of k's </b></font>
+
 * *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*

### 2017-06-18 MSDS7333-baldree-case6-fx.r
finished up part 1

@@ -350,15 +350,15 @@ plotSSErrors = function(err, K){
   oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
   par(mai=c(.8, .8, .4, .25))
 
-  plot(y = err, x = (1:K),  type = "l", lwd= 2, ylim = c(1200, 2100), cex.main=.8, cex.axis=.8, cex.lab=.8,
+  plot(y = err, x = (1:K),  type = "l", lwd= 2, ylim = c(900, 2100), cex.main=.8, cex.axis=.8, cex.lab=.8,
        xlab = "Number of Neighbors",
-       ylab = "Sum of Square Errors",
+       ylab = "Sum of Squared Errors",
        main = "Cross Validation Section of k")
 
   rmseMin = min(err)
   kMin = which(err == rmseMin)[1]
   segments(x0 = 0, x1 = kMin, y0 = rmseMin, col = gray(0.4), lty = 2, lwd = 2)
-  segments(x0 = kMin, x1 = kMin, y0 = 1100,  y1 = rmseMin, col = grey(0.4), lty = 2, lwd = 2)
+  segments(x0 = kMin, x1 = kMin, y0 = 800,  y1 = rmseMin, col = grey(0.4), lty = 2, lwd = 2)
   text(x = kMin - 2, y = rmseMin + 40, label = as.character(round(rmseMin)), col = grey(0.4), cex = .6)
 
   par(oldPar)

### 2017-06-18 MSDS7333-baldree-case6.r

@@ -11,7 +11,7 @@
 #   Which MAC address yields the best prediction of location?
 #   Does using data for both MAC addresses simultaneously yield more, or less, accurate prediction of location?
 #
-# 2. Implement alternative k-nearest precition method using weights on received signal strength.
+# 2. Implement alternative k-nearest prediction method using weights on received signal strength.
 #   For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?
 #     Use calcError() to compare approaches.
 #
@@ -20,9 +20,8 @@
 
 # two digit precision for printing and formatting
 options(digits = 2)
-options(error = recover, warn = 2)
+options(error = recover, warn = 1)
 # include necessary functions
-#if (!exists("processLine", mode="function")) source("MSDS7333-baldree-case6-fx.r")
 source("MSDS7333-baldree-case6-fx.r", print.eval=TRUE)
 
 #### Part 1: Access point MAC address analysis of 00:0f:a3:39:e1:c0 and 00:0f:a3:39:dd:cd ####
@@ -116,19 +115,9 @@ names(onlineSummary)
 # what if we didn't want to collapse the signal strengths across the m angles, and
 # instead return a set of mx166 signals for each access point?
 
-
-for (ap in c("", rejectedAP, chosenAP)){
-  test = onlineSummary[, !(names(onlineSummary) %in% ap)]
-  train = subset(offlineSummary, mac != ap)
-  err = predictionErrors(test, train, numberOfAngles = 1, k = 3)
-  print(paste(ap, err))
-}
-
-
 # iterate through 1 to 3 neighbors with 1 to 3 angles for each scenario to determine best combination
 # for cross validation
 # result: it is evident that we should include both access points for location predictions.
-
 for (ap in c("None", rejectedAP, chosenAP)){
   if (ap == "None"){
     test = onlineSummary
@@ -149,4 +138,47 @@ for (ap in c("None", rejectedAP, chosenAP)){
 }
 
 # perform cross validation and plot for desired AP for a range of k's
+# for each fold, we want to find k-NN estimates from 1..K and aggregate errros over the v folds.
+# choose a v-fold of 11
+v = 11
+# number of locations
+permuteLocs = sample(unique(offlineSummary$posXY))
+# matrix locations into folds
[…]

### 2017-06-18 MSDS7333-baldree-case6-fx.r
fixed bug

@@ -295,7 +295,8 @@ findNN = function(newSignal, trainingSubset) {
   # Returns:
   #   training neighbors
   #
-  diffs = apply(trainingSubset[ , 4:9], 1, function(x) x - newSignal)
+  cols = length(trainingSubset)
+  diffs = apply(trainingSubset[ , 4:cols], 1, function(x) x - newSignal)
   dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) )
   closest = order(dists)
   return(trainingSubset[closest, 1:3 ])
@@ -325,6 +326,7 @@ predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3){
   # could also use a different metric besides Euclidean like Manhattan.
   # could use medians instead of averages when combining neighbors if the distribution of values are quite skewed.
   estXY = lapply(closeXY, function(x) sapply(x[ , 2:3], function(x) mean(x[1:k])))
+
   return(do.call("rbind", estXY))
 }
 

### 2017-06-18 MSDS7333-baldree-case6.r

@@ -122,12 +122,14 @@ for (ap in c("None", rejectedAP, chosenAP)){
   if (ap == "None"){
     test = onlineSummary
     train = offlineSummary
+    newSignals = test[ , 6:12]
   } else {
     test = onlineSummary[, !(names(onlineSummary) %in% ap)]
     train = subset(offlineSummary, mac != ap)
+    newSignals = test[ , 6:11]
   }
 
-  estXY = predXY(newSignals = test[ , 6:11],
+  estXY = predXY(newSignals = newSignals,
                  newAngles = test[ , 4],
                  train,
                  numAngles = 3, k = 3)
@@ -163,7 +165,7 @@ for (j in 1:v) {
 
   # loop through neighbors
   for (k in 1:K) {
-    estFold = predXY(newSignals = testFold[ , 6:11],
+    estFold = predXY(newSignals = testFold[ , 6:12],
                      newAngles = testFold[ , 4],
                      trainFold, numAngles = 3, k = k)
     err[k] = err[k] + calcError(estFold, actFold)
@@ -174,11 +176,12 @@ plotSSErrors(err, K)
 
 # experiment with a few k's based on above analysis to determine k
 # that results in lowest sum squared error on test data.
-estXY = predXY(newSignals = onlineSummary[ , 6:11],
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
                newAngles = onlineSummary[ , 4],
-               offlineSummary, numAngles = 3, k = 4)
+               offlineSummary, numAngles = 3, k = 5)
 actXY = onlineSummary[ , c("posX", "posY")]
 calcError(estXY, actXY)
 
 #### Part 2: Alternatvie k-nearest ####
+
 

### 2017-06-18 MSDS7333-baldree-case6-fx.r
added a candidate part 2.

@@ -298,11 +298,13 @@ findNN = function(newSignal, trainingSubset) {
   cols = length(trainingSubset)
   diffs = apply(trainingSubset[ , 4:cols], 1, function(x) x - newSignal)
   dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) )
-  closest = order(dists)
-  return(trainingSubset[closest, 1:3 ])
+  #closest = order(dists)
+  closest = cbind(trainingSubset[, 1:3 ], dists)
+  return(closest[order(dists),])
+  #return(trainingSubset[closest, 1:3 ])
 }
 
-predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3){
+predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3, weighted=FALSE){
   # Predict the XY coordinates given a list of signals along with their angles measured and training data.
   # k neighbors will be found to make the prediction of coordinate.
   #
@@ -312,6 +314,8 @@ predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3){
   #   trainingData: data needed for k-nn
   #   numAngles: angles we want to use for prediction
   #   k: number of neighbors to use for prediction
+  #   weighted: if TRUE, distances to neighbors will be used to weight XY for each neighbor to determine
+  #             location.
   #
   closeXY = list(length = nrow(newSignals))
 
@@ -325,8 +329,17 @@ predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3){
   # are 1/d / sum(1/d).
   # could also use a different metric besides Euclidean like Manhattan.
   # could use medians instead of averages when combining neighbors if the distribution of values are quite skewed.
-  estXY = lapply(closeXY, function(x) sapply(x[ , 2:3], function(x) mean(x[1:k])))
-
+  if (!weighted) {
+    estXY = lapply(closeXY, function(x) sapply(x[ , 2:3], function(x) mean(x[1:k])))
+  } else {
+    # sum X and Y weighted neighbor values.
+    estXY = lapply(closeXY, function(x) {
+      wt = (1/x[1:k, 4])/sum(1/x[1:k, 4])
+      xy = cbind(x[1:k,], wt)
+      # sum weighted neighbors
+      return(colSums(xy[1:4, 2:3] * xy[1:4, 4]))
+    })
+  }
   return(do.call("rbind", estXY))
 }
 

### 2017-06-18 MSDS7333-baldree-case6.r

@@ -184,4 +184,12 @@ calcError(estXY, actXY)
 
 #### Part 2: Alternatvie k-nearest ####
 
+# create weights for neighbors based on signal strength and multiple times X and Y and sum neighbors
+# to get X and Y.
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
+               newAngles = onlineSummary[ , 4],
+               offlineSummary, numAngles = 3, k = 5, weighted=TRUE)
+actXY = onlineSummary[ , c("posX", "posY")]
+calcError(estXY, actXY)
+
 

### 2017-06-18 MSDS7333-baldree-case6-fx.r
Add wt_signals() method to MSDS7333-baldree-case6-fx.R file
Add wt_signals() method to MSDS7333-baldree-case6-fx.R file.  This was
recommended by Prof Holt.

@@ -379,3 +379,25 @@ plotSSErrors = function(err, K){
   par(oldPar)
 }
 
+wt_signals <- function(ByAvgSig){
+  
+  # Name: wt_signals
+  # Create weighted data
+  #
+  # Args:
+  #   ByAvgSig:    estimate XY line segment
+  #   
+  # Returns:
+  #   log_ang by weighted_signals matrix
+  
+  # calculate signals
+  signals <- select(ByAvgSig, -posXY, -posX, -posY, -angle) 
+  
+  # calculate log_ang
+  loc_ang <- select(ByAvgSig, posXY, posX, posY, angle)
+  
+  # calculate weighted_signals 
+  weighted_signals <- (1/signals) / rowSums(1/signals)
+  
+  return(cbind(loc_ang, weighted_signals))
+}

### 2017-06-18 MSDS7333-baldree-case6-fx.r
Update function preamble for wt_signals.
Update function preamble for wt_signals.

@@ -383,7 +383,17 @@ wt_signals <- function(ByAvgSig){
   
   # Name: wt_signals
   # Create weighted data
+  # Weighted Average Predictor model, where the weights are inversely proportional to the 
+  # "distance" (in signal strength) from the test observation.
+  # This allows us to include the k points that are close, but to differentiate between them
+  # by how close they actually are. 
   #
+  # The weights might be
+  #
+  #         (1/di)/ from 1 to K SUM(1/di)
+  #
+  # for the i-th closest neighboring observation where di is the distance from our new test
+  # point to this reference point (in signal strength space).
   # Args:
   #   ByAvgSig:    estimate XY line segment
   #   

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Still running into issues.
Will investigate tomorrow.
Added rest of code

@@
 date: "June 17, 2017"
-output: html_notebook
+output:
+  html_notebook: default
+  html_document: default
 ---
@@
 ## Introduction
+(insert text here)
 
@@
 ```{r}
+
+```
+
+```{r}
 df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
@@
 ```
-<font color = "red"> &uarr; Having some trouble displaying the individual graphs here. &uarr; </font>
+<font color = "red"> <b> &uarr; having issues with this diplaying properly... &uarr;</b> </font>
 
@@
 ```{r}
-for (ap in c("", rejectedAP, chosenAP)){
-  test = onlineSummary[, !(names(onlineSummary) %in% ap)]
-  train = subset(offlineSummary, mac != ap)
-  err = predictionErrors(test, train, numberOfAnglers = 1, k = 3)
-  print(paste(ap, err))
-}
+source("MSDS7333-baldree-case6-fx.r", print.eval = TRUE)
 ```
@@
     train = offlineSummary
+    newSignals = test[ , 6:12]
   } else {
     test = onlineSummary[, !(names(onlineSummary) %in% ap)]
+    train = subset(offlineSummary, mac != ap)
+    newSignals = test[,6:11]
   }
 
-  estXY = predXY(newSignals = test[, 6:11],
+  estXY = predXY(newSignals = newSignals,
                  newAngles = test[, 4],
@@
                  numAngles = 3, k = 3)
-  actXY = test[, c("posX", "posY")]
+  actXY = test[ , c("posX", "posY")]
   err = calcError(estXY, actXY)
@@
 
-* *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
-* *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*
+Perform cross validation and plot for desired AP for a range of k's for each fold. We want to find k-NN estimates from 1...k and aggregate errors over the v folds.
 
-### Part 2
+Choose a v-fold of 11
+```{r}
+v = 11
+```
+
+```{r}
+
+# number of locations
+permuteLocs = sample(unique(offlineSummary$posXY))
+
+```
+
+```{r}
+# matrix locations into folds
+permuteLocs = matrix(permuteLocs, ncol = v, nrow = floor(length(permuteLocs)/v))
+```
+<font color = "Red"><b>&uarr; come back to this one &uarr;</b></font>
+
+```{r}
+# summarize and format offline
+keepVars = c("posXY", "posX", "posY", "orientation", "angle")
+onlineCVSummary = reshapeSS(offline, keepVars = keepVars, sampleAngle = TRUE)
+```
+
+```{r}
+# number of k's
+K = 10
+# errors array
+err = rep(0, K)
+
+# loop through folds
+for (j in 1:v) {
+  testFold = subset(onlineCVSummary, posXY %in% permuteLocs[ , j])
+  trainFold = subset(offlineSummary, posXY %in% permuteLocs[ , -j])
+  actFold = testFold[ , c("posX", "posY")]
+
+  # loop through neighbors
+  for (k in 1:K) {
+    estFold = predXY(newSignals = testFold[ , 6:12],
+                     newAngles = testFold[ , 4],
+                     trainFold, numAngles = 3, k = k)
+    err[k] = err[k] + calcError(estFold, actFold)
+  }
+}
+# plot k to sum of squared errors
+plotSSErrors(err, K)
+```
+
+```{r}
+# experiment with a few k's based on above analysis to determine k
+# that results in lowest sum squared error on test data.
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
+               newAngles = onlineSummary[ , 4],
+               offlineSummary, numAngles = 3, k = 5)
+actXY = onlineSummary[ , c("posX", "posY")]
+calcError(estXY, actXY)
+```
 
 *** 
+#Part 2: Alternatvie k-nearest neighbors
+```{r}
+# create weights for neighbors based on signal strength and multiple times X and Y and sum neighbors
+# to get X and Y.
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
+               newAngles = onlineSummary[ , 4],
+               offlineSummary, numAngles = 3, k = 5, weighted=TRUE)
+actXY = onlineSummary[ , c("posX", "posY")]
+calcError(estXY, actXY)
+```
 
-### Preliminary Functions to be used: 
-## Let's break this section up and explain each function individually
-
-## The Output Achieved (this should be included as close as possible to the output or figures as possible)
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+***

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Created and filled out the intro. Needs revision

@@
 * *Use calcError() to compare approaches*
+<font color = "green">
 
 ## Introduction
-(insert text here)
+#### *Explain the case study*
+- Discuss the goal of the case study
 
+For this case study, the overall goal is to use K-Nearest Neighbors (kNN) and an alternative weighted method to be able to determine how different handheld devices moved through a building <font color = "red">(at Dartmouth.
+Let's confirm this and add some more detail)</font>.
+This can be achieved with an analysis of how the different Media Access Control (MAC) addresses for each of the devices interacted with the Access Points (APs) of the building.
+In this case, the APs are routers with known locations throughout the building.
+
+- Discuss the different Access points and MAC Addresses
+
+The access points as stated above allow for the tracking of the devices, thanks to the history of times, signal strengths and angles recorded for each MAC address with respect to each access point.<font color = "red">&larr;(consider fragmenting this sentence)</font> From the information that has been given, there are known to be six APs on the building floor in question.
+However, this is a problem since there are seven APs recorded.
+To determine which AP to exclude from the analysis, we will need to compare these two addresses:
+
+    chosenAP   = "00:0f:a3:39:e1:c0"
+    rejectedAP = "00:0f:a3:39:dd:cd"
+
+and decide which to exclude and why.
+
+  - Discuss why we're trying to remove one
+
+Removing one of these APs will allow us to remove noise from the resulting model.
+Since only six APs are recorded as existing on the floor, the presence of a seventh points to external interference.
+Once the superfluous address has been identified, we will execute testing to determine whether there will be any benefit added to the model by keeping or excluding said data.
+This decision will ultimately lead to a better prediction of location of each device within this floorplan.
+
+- BRIEF overview of K-Nearest Neighbors
+
+K-Nearest Neighbors (kNN) is an ideal method to use in this case as the goal is to be able to predict the path of each device on their journey through the building.
+To utilize this method, we use separate testing and training data.
+The training data are known observations from well documented locations of both the handheld devices and APs.
+These locations are represented by the angle at which the AP received the signal from the device as well as the overall signal strength received.
+
+With this training data, each observation is evaluated along with the observations which occur in the general vicinity in the training data.
+These occurrences allow for the different probabilities of occurrence to be proposed through the model.
+Once this model has been generated for k neighbors, it is then tested against the unobserved test data and adapted to suit what actually occurs.
+This allows for an accurate gauge of accuracy of the model and is recreated for a span of possible k neighbors to exhaust the total possible number of observations.
+
+- Discuss the alternative weighted method.
+
+In contrast to the kNN method, another option to explore is a weighted method based on the signal strength received at the access point.
+This method involves defining a new metric where the weight of each observation is inversely proportional to the distance from the test observation.
+From a topographical standpoint, the observations which are closest to the observation will have a much higher influence than those which exist further away.
+
+- Discuss potential for prediction and what we expect to come of this.
+
+The evaluation of this data will allow us to determine which data will be critical to the creation of an accurate model and which will not.
+Since we have six access points and seven addresses, this thorough investigation will help us decide which address is simply an external interference and if it is beneficial to keep involving said data.
+As a result, we expect to have a more thorough understanding of the difference between k-Nearest Neighbors and the weighted alternative.
+
+</font>
 ***

### 2017-06-18 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
more updates
adding to the Approach Section

@@
 This can be achieved with an analysis of how the different Media Access Control (MAC) addresses for each of the devices interacted with the Access Points (APs) of the building.
-In this case, the APs are routers with known locations throughout the building.
+In this case, the APs are routers with static locations throughout the building.
 
@@
 
-The access points as stated above allow for the tracking of the devices, thanks to the history of times, signal strengths and angles recorded for each MAC address with respect to each access point.<font color = "red">&larr;(consider fragmenting this sentence)</font> From the information that has been given, there are known to be six APs on the building floor in question.
+The access points as stated above allow for the tracking of the devices thanks to the logged history of times, signal strengths and angles for each MAC address with respect to each access point.<font color = "red">&larr;(consider fragmenting this sentence)</font> From the information that has been given, there are known to be six APs on the building floor in question.
 However, this is a problem since there are seven APs recorded.
@@
 </font>
+
 ***
@@
 
+To fully answer this question, we will need to explore exactly what is happening with each MAC address. The results of this exploration will define the parameters needed for such a decision.
+
 #### Here we will create a Signal Collection Map
@@
 Plotting the individual signal strength distributions for each AP angle will help add perspective to the information we are working with.
-For the sake of consistency, we will use the same stationary coordinates listen in the course book (2, 12) (Add reference here.)
+For the sake of consistency, we will use the same stationary coordinates listed in the course book (coord : (2, 12)) <font color = "red">(Add reference here.)</font>
 
@@
 From the literature, we know the higher the decibel-milliwatts means a stronger signal strength.
-<font color="red">**(consider revising this sentence.
-Seems incomplete.)**</font> A stronger signal strength indicates a closer position to the AP since signal strengths decrease exponentially with distance.
-Comparing the signal strength distributions in this chart, we see "e1:c0" has a stronger signal strength at this location than "dd:cd."
+A stronger signal strength indicates a closer physical position to the AP since signal strengths decrease exponentially with distance.<font color="red">Find this in the book and cite it.</font> Comparing the signal strength distributions in this chart, we see "e1:c0" has a stronger signal strength at this location than "dd:cd."
 
-What if we compared the two addresses from the opposite end of the building? 
+What if we compared the two addresses from the opposite side of the building? 
 ```{r}
@@
 #### Time to investigate Signal Density
-To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each angle.
+To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each 45 degree angle.
 This will help display the defining characteristics for each address.
-
-```{r}
-
-```
+The location used for this analysis is the (24,4) from the book.
 
@@
 <font color = "red"> <b> &uarr; having issues with this diplaying properly... &uarr;</b> </font>
+
+From this plot, it can be seen that the rejectedAP has a higher signal density at lower strength while the acceptedAP has a higher density coverage at a higher strength.
+Judging by the distributions themselves, it seems as if many of the plots are normally distributed while some of the other plots (i.e.
+"dd:cd" at 180 and "e1:c0" at 315) are heavily skewed or even bimodal.
+Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle.
+This would suggest the median of each distribution varies by angle and attributes to a difference in perceived strength.
+
+This evidence would suggest both addresses are kept within the analysis to increase the coverage for the entire spectrum of strengths.
 

### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
pushing updates

@@
 
-```{r}
+```{r, fig.height=10}
 df = subset(offline, posX == 24 & posY == 4 & !(mac %in% remainingAP), c(signal, angle, mac))
@@
 ```
-<font color = "red"> <b> &uarr; having issues with this diplaying properly... &uarr;</b> </font>
 
@@
 ```{r}
-offline$posXY = paste(offline$posX, offline$posY, sep="-")
+offlineSummary = createOfflineSummary(offline)
+```
 
+### <font color = "red"> &darr; this section may no longer be needed. Now missing from the original code. &darr;</font>  
+```{r}
+#offline$posXY = paste(offline$posX, offline$posY, sep="-")
+#
 # create data frames for each combination
-byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
+#byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
 ```
@@
 ```{r}
-signalSummary = lapply(byLocAngleAP, function(oneLoc) {
-  ans = oneLoc[1, ]
-  ans$medSignal = median(oneLoc$signal)
-  ans$avgSignal = mean(oneLoc$signal)
-  ans$num = length(oneLoc$signal)
-  ans$sdSignal = sd(oneLoc$signal)
-  ans$iqrSignal = IQR(oneLoc$signal)
-  ans
-})
-offlineSummary = do.call("rbind", signalSummary)
+#signalSummary = lapply(byLocAngleAP, function(oneLoc) {
+#  ans = oneLoc[1, ]
+#  ans$medSignal = median(oneLoc$signal)
+#  ans$avgSignal = mean(oneLoc$signal)
+#  ans$num = length(oneLoc$signal)
+#  ans$sdSignal = sd(oneLoc$signal)
+#  ans$iqrSignal = IQR(oneLoc$signal)
+#  ans
+#})
+#offlineSummary = do.call("rbind", signalSummary)
 ```
+### <font color = "red"> &uarr; this section may no longer be needed. Now missing from the original code. &uarr;</font>
 
-The Standard Deviation of the average signal strength per targeted AP
+The Standard Deviation of the average signal strength per targeted AP:
 
@@
 
+From this figure, it is apparent that there is considerable overlap between the two distributions of signal strength.
+This is evidence in favor of keeping both addresses in the analysis to provide greater coverage.
+
 #### Let's dive into the K-Nearest Neighbors:
@@
 macs = unique(offlineSummary$mac)
+macs
 online = readData("online.final.trace.txt", subMacs = macs)

### 2017-06-19 MSDS7333-baldree-case6-fx.r
update on part 2

@@ -283,28 +283,39 @@ selectTraininingData = function(newObservationAngle, df = NULL, m = 1){
   return(reshapeSS(subset, varSignal = "avgSignal"))
 }
 
-findNN = function(newSignal, trainingSubset) {
-  # want to look the distance in terms of signal strengths from these training data to the new
-  # data point. we need to calculate teh distrance from the new point to all observations in the
+findNN = function(newSignal, trainingSubset, k, exp=.5) {
+  # we need to calculate the distrance from the new point to all observations in the
   # training set with findNN().
-  # returns locations of the training observations in order of closeness to the new observation's signal strength.
+  # returns locations of the training observations in order of closeness to the new
+  # observation's signal strength.
+  #
+  # With inspiration from Andrew Clark.
   #
   # Args:
   #   newSignal: signal of new observation
   #   trainingSubset: training data to find neighbors
+  #   k: neighbors
+  #   exp: the exponential to apply to the squared distance. the default is .5 (Euclidean).
   # Returns:
   #   training neighbors
   #
   cols = length(trainingSubset)
   diffs = apply(trainingSubset[ , 4:cols], 1, function(x) x - newSignal)
-  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) )
-  #closest = order(dists)
-  closest = cbind(trainingSubset[, 1:3 ], dists)
-  return(closest[order(dists),])
-  #return(trainingSubset[closest, 1:3 ])
+  dists = apply(diffs, 2, function(x) (sum(x^2)^exp))
+  closest = order(dists)
+
+  trainingSubset$wts = (1/dists[closest[1]])/sum(1/dists[closest[1:k]])
+  trainingSubset$wtPosX = trainingSubset$wts * trainingSubset$posX
+  trainingSubset$wtPosY = trainingSubset$wts * trainingSubset$posY
+
+  if (length(trainingSubset)==13) {
+    return(trainingSubset[closest, c(1:3, 12:13) ])
+  } else {
+    return(trainingSubset[closest, c(1:3, 11:12) ])
+  }
 }
 
-predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3, weighted=FALSE){
[…]

### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
udpates

@@
 
+First and foremost, let's display the unique MAC addresses: 
 ```{r}
@@
 macs
+```
+
+```{r}
 online = readData("online.final.trace.txt", subMacs = macs)
@@
 ```
+This output shows the signal strengths at each angle. This will be useful later!
+<font color = "red"><b>&uarr;Explain why &uarr;</b></font>
 
 Organize the data where each AP is in a column. 
-
 ```{r}
@@
 ```
-
-<font color = "red">(Continued from baldree-case6.r)</font>
 
@@
 
+When it comes to the kNN, we will need to include the data which include the angles close to the point in question.
+As observed before, the angle matters!
+For this exercise, we will need to abide by the following outlines:
+
+  * If one angle is desired, then we will include angles which match the rounded orientation of the new observation  
+  * If we want two angles, two multiples of 45 degrees will be chosen in order to encapsulate the nw observation's orientation  
+  * For three angles, the nearest 45 degree increment is chosen and either side will be used.
+
+What if we didn't want to collapse the signal strengths across the *m* angles and, instead, we return a set of mx166 signals for each access point?
+
 ```{r}
@@
 
+Here, we begin by iterating through 1-3 neighbors at 1-3 angles for each combination to determine which will be best suited for cross-validation:
+
 ```{r}
-#Iterate through 1 to 3 neighbors with 1 to 3 angles for each scenario to determine best combination
-#for cross validation
 for (ap in c("None", rejectedAP, chosenAP)){
@@
 
-Perform cross validation and plot for desired AP for a range of k's for each fold. We want to find k-NN estimates from 1...k and aggregate errors over the v folds.
+Perform cross validation and plot for desired AP for a range of k's for each fold. We want to find k-NN estimates from 1 to k and aggregate errors over the v folds.
 
@@
 ```{r}
-
 # number of locations
@@
 ```
-<font color = "Red"><b>&uarr; come back to this one &uarr;</b></font>
+<font color = "Red"><b>&uarr; come back to this one &uarr; For some reason it only runs after it throws an error (e.g.change matrix to sapply, run, change back and run again</b></font>
 

### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added meat to the bones

@@
 
-### Assignment: 
-1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both: 
-    + chosenAP = "00:0f:a3:39:e1:c0"
-    + rejectedAP = "00:0f:a3:39:dd:cd"
-* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
-* *Which MAC address yields the best prediction of location?including determining locations by using data corresponding to both MAC addresses*
-* *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*
+### Table of Contents:
+<li>[Assignment Requirements](#AssignmentOutline)</li>
+<li>[Introduction](#Introduction)</li>
+<li>[Part 1](#Part1)</li>
+  <li>[Signal Collection](#SigCol)</li>
+  <li>[Signal Strength](#SigStren)</li>
+  <li>[Signal Density](#SigDen)</li>
+  <li>[Discussion](#Discussion)</li>
+<li>[Part2](#Part 2)</li>
+
+<li>[References](#References)</li>
+<li>[Appendix](#Appendix)</li>
+
+<a id='AssignmentOutline'>
+
+### <b>Assignment</b> </a>
+1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both the chosen and rejected Access Points:
+    + *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
+    + *Which MAC address yields the best prediction of location? (Including determining locations by using data corresponding to both MAC addresses)*
+    + *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*
+
+2. Implement alternative k-nearest precition method using weights on received signal strength.
+    + *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
+    + *Use calcError() to compare approaches*
 
@@
 
-2. Implement alternative k-nearest precition method using weights on received signal strength.
-* *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
-* *Use calcError() to compare approaches*
-<font color = "green">
+<a id='Introduction'>
 
-## Introduction
+### <b>Introduction</b></a>  
 #### *Explain the case study*
+
+<font color = "purple"><b>>>>>Everything in this color to be removed at final review<<<<</b></font>
+
+<font color = "purple"><b>
 - Discuss the goal of the case study
+</b></font>
 
@@
 
+<font color = "purple"><b>
 - Discuss the different Access points and MAC Addresses
+</b></font>
 
@@
 
+<font color = "purple"><b>
 - Discuss why we're trying to remove one
+</b></font>
 
@@
 
+<font color = "purple"><b>
 - BRIEF overview of K-Nearest Neighbors
+</b></font>
 
@@
 
+<font color = "purple"><b>
 - Discuss the alternative weighted method. 
+</b></font>
 
@@
 
+<font color = "purple"><b>
 - Discuss potential for prediction and what we expect to come of this.
+</b></font>
 
@@
 
-</font>
-
 ***
 
-## The Approach (Explained)
-
 ### Part 1
+#### The Approach (Explained)
 
@@
 ```
+
 Before continuing, we will need to address another file which houses the predefined functions. This helps to save notebook real estate and prevents confusion:
@@
 
-* *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
+#### * *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
 
@@
 ```
-
 From this plot, it can be seen that the rejectedAP has a higher signal density at lower strength while the acceptedAP has a higher density coverage at a higher strength.
@@
 
-This evidence would suggest both addresses are kept within the analysis to increase the coverage for the entire spectrum of strengths.
+***
 
@@
 
-### <font color = "red"> &darr; this section may no longer be needed. Now missing from the original code. &darr;</font>  
 ```{r}
-#offline$posXY = paste(offline$posX, offline$posY, sep="-")
-#
+offline$posXY = paste(offline$posX, offline$posY, sep="-")
+
 # create data frames for each combination
-#byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
+byLocAngleAP = with(offline, by(offline, list(posXY, angle, mac), function(x) x))
 ```
@@
 ```{r}
-#signalSummary = lapply(byLocAngleAP, function(oneLoc) {
-#  ans = oneLoc[1, ]
-#  ans$medSignal = median(oneLoc$signal)
-#  ans$avgSignal = mean(oneLoc$signal)
-#  ans$num = length(oneLoc$signal)
-#  ans$sdSignal = sd(oneLoc$signal)
-#  ans$iqrSignal = IQR(oneLoc$signal)
-#  ans
-#})
-#offlineSummary = do.call("rbind", signalSummary)
+signalSummary = lapply(byLocAngleAP, function(oneLoc) {
+  ans = oneLoc[1, ]
+  ans$medSignal = median(oneLoc$signal)
+  ans$avgSignal = mean(oneLoc$signal)
+  ans$num = length(oneLoc$signal)
+  ans$sdSignal = sd(oneLoc$signal)
+  ans$iqrSignal = IQR(oneLoc$signal)
+  ans
+})
+offlineSummary = do.call("rbind", signalSummary)
 ```
-### <font color = "red"> &uarr; this section may no longer be needed. Now missing from the original code. &uarr;</font>
 
@@
 This is evidence in favor of keeping both addresses in the analysis to provide greater coverage.
+
+***
+
+#### We now have enough information to address these questions before an analysis of kNN is conducted:
+
+* **Which MAC address yields the best prediction of location? (Include determining locations by using data corresponding to both MAC addresses)**
+* **Does using data for both MAC addresses simultaneously yield a more or less accurate prediction of location?**
+
+The evidence would suggest both addresses should be kept within the analysis to increase the coverage for the entire spectrum of strengths.
+The address chosen to be rejected not only has a lower signal strength when measured from either end of the building, it also shows to have a higher signal density when it is further away from the receiver.
+The address chosen to be kept has a higher signal density closer to the receivers.
+By keeping both addresses, the model should be able to generate a more accurate prediction of the layout of the building.
+This theory will be tested as we tour through k-Nearest Neighbors.
 
@@
 What if we didn't want to collapse the signal strengths across the *m* angles and, instead, we return a set of mx166 signals for each access point?
-
-```{r}
-source("MSDS7333-baldree-case6-fx.r", print.eval = TRUE)
-```
 
@@
 *** 
-#Part 2: Alternatvie k-nearest neighbors
+<a id='Part2'>
+
+#Part 2: Alternatvie k-Nearest Neighbors
+</a>
 ```{r}
@@
 ***
+<a id='References'>
+
+### References  
+</a>
+
+<a id='Appendix'>
+
+### Appendix (including function code)  
+</a>
+This is the code used to define the functions used within the study. It can be found under 'MSDS7333-baldree-case6-fx.r' which is included in the final submission  
+```{r}
+processLine = function(x)
+{
+  # Process a line of text returning a 11x10 matrix of tokens
+  #
+  # Args:
+  #   x: The line of text
+  #
+  tokens = strsplit(x, "[;=,]")[[1]]
+
+  # return NULL if only 10 tokens
+  if (length(tokens) == 10)
+    return(NULL)
+
+  tmp = matrix(tokens[-(1:10)], ncol = 4, byrow = TRUE)
+  mat = cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow(tmp), 6, byrow = TRUE), tmp)
+}
+
+roundOrientation = function(angles) {
+  # Round an array of degree angles to the nearest 45 degree angle.
+  #
+  # Args:
+  #   angles: The line of text
+  #
+  refs = seq(0, by = 45, length  = 9)
+  q = sapply(angles, function(o) which.min(abs(o - refs)))
+  c(refs[1:8], 0)[q]
+}
+
+readData = function(filename = 'offline.final.trace.txt',
+                    subMacs = c("00:0f:a3:39:e1:c0", "00:0f:a3:39:dd:cd", "00:14:bf:b1:97:8a",
+                                "00:14:bf:3b:c7:c6", "00:14:bf:b1:97:90", "00:14:bf:b1:97:8d",
+                                "00:14:bf:b1:97:81")) {
+  # Process the csv file peforming the following steps:
+  #   Read all non-comment lines.
+  #   Call processLine() for each line of text.
+  #   Convert list of text into a data frame
+  #   Name columns of data frame
+  #   Define column types
+  #   Convert millisecond time to seconds
+  #   Round angle orientations
+  #   Drop data
+  #
+  # Args:
+  #   filename: The name of the csv file to process. The default is 'offline.final.trace.txt.'
+  #   subMacs: The list of access point MAC addresses to filter the dataset down to. A default list is provided.
+  #
+  # Returns:
+  #   The data frame
+  #
+
+  # process csv into matrix of tokens
+  txt = readLines(filename)
+  lines = txt[substr(txt, 1, 1) != "#"]
+  tmp = lapply(lines, processLine)
+  # convert into tokens into a data frame
+  df = as.data.frame(do.call("rbind", tmp), stringsAsFactors = FALSE)
+
+  # variable names
+  names(df) = c("time", "scanMac", "posX", "posY", "posZ", "orientation", "mac", "signal", "channel", "type")
+
+  # convert numeric values
+  numVars = c("time", "posX", "posY", "orientation", "signal")
+  df[numVars] = lapply(df[numVars], as.numeric)
+
+  # convert time to POSIX
+  df$rawTime = df$time
+  df$time = df$time / 1000
+  class(df$time) = c("POSIXt", "POSIXct")
+
+  # round orientations to nearest 45
+  df$angle = roundOrientation(df$orientation)
+
+  # keep only signals from access points
+  df = df[df$type == "3",]
+  df = df[, "type" != names(df)]
+
+  # drop scanMac, posZ, channel, and type - no info in them
+  df = df[, !(names(df) %in% c("scanMac", "posZ"))]
+
+  # drop more unwanted access points
+  df = df[df$mac %in% subMacs,]
+
+  # drop channel
+  df = df[, "channel" != names(df)]
+
+  return(df)
+}
+
+plotBoxplotSignalStrength = function(df) {
+  # Box plot signal strength
+  #
+  # Args:
+  #   df: data frame of data to analyze
+  #
+  oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
+  par(mai=c(.8, .8, .5, .25))
+  library (lattice)
+
+  print(bwplot(signal ~ factor(angle) | mac, data = df, layout = c(2,1),
+               main="Signal Strength Distribution for Access Points",
+               xlab="Angle Measured (deg)", ylab="Signal Strengh (dBM)"))
+  par(oldPar)
+}
+
+plotDensitySignalStrength = function(df) {
+  # Density plot signal strength
+  #
+  # Args:
+  #   df: data frame of data to analyze
+  #
+  oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
+  par(mai=c(.8, .8, .5, .25))
+  library (lattice)
+
+  print(densityplot( ~ signal | mac + factor(angle), data = df, bw = 0.5, plot.points = FALSE,
+        main="Signal Strength Distribution for Access Points",
+        xlab="Angle Measured (deg)", ylab="Signal Strengh (dBM)"))
+  par(oldPar)
+}
+
+plotStdDevSignalStrength = function(df) {
+  # Standard Deviation of signal strength average
+  #
+  # Args:
+  #   df: data frame of data to analyze
+  #
+  oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
+  par(mai=c(.8, .8, .5, .25))
+  library (lattice)
+  breaks = seq(-90, -30, by=5)
+
+  print(bwplot(sdSignal ~ cut(avgSignal, breaks=breaks) | mac, data=df,
+               xlab = "Mean Signal (dBM)", ylab = "Signal Standard Deviation (dBM)",
+               main = "Standard Deviation of Average Signal Strength from Detector to Access Point",
+               cex.main=.8, cex.axis=.8, cex.lab=.8, layout = c(1, 2)))
+  par(oldPar)
+}
+
+plotSignalMaps = function(macAddresses) {
+  # Plot the number of signals per XY location at XY location.
+  #
+  # Args:
+  #   macAddresses: a list of macs to plot signals
+  #
+  oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
+  par(mfrow = c(2, 1), mai=c(.8, .8, .4, .25))
+
+  for (i in macAddresses) {
+    # filter data set
+    df = offline[offline$mac %in% i, ]
+    # location of datapoints filtered
+    locDF = with(df, by(df, list(posX, posY), function(x) x))
+    # drop locations that were not observed, null
+    locDF = locDF[!sapply(locDF, is.null)]
+
+    # determine the number of observations recorded at each location
+    locCounts = sapply(locDF, nrow)
+
+    # keep position information with location
+    locCounts = sapply(locDF, function(df) c(df[1, c("posX", "posY")], count = nrow(df)))
+
+    # transpose matrix
+    locCounts = t(locCounts)
+    plot(locCounts, type = "n", xlab = "Position X", ylab = "Position Y",
+         main = paste("Total Signals from Detector to Access Point: ", i),
+         cex.main=.8, cex.axis=.8, cex.lab=.8)
+    text(locCounts, labels = locCounts[, 3], cex = .6, srt = 45)
+  }
+  par(oldPar)
+}
+
+castOnline = function(df) {
+  # Cast dataframe of online data so each access point (AP) is in its own column.
+  # Matrix is now a 1x7 since we are including rejected AP.
+  #
+  # Args:
+  #   df: online data frame
+  #
+  keepVars = c("posXY", "posX","posY", "orientation", "angle")
+  numOfAPs = length(unique(df$mac))
+  byLoc = with(df,
+               by(df, list(posXY),
+                  function(x) {
+                    ans = x[1, keepVars]
+                    avgSS = tapply(x$signal, x$mac, mean)
+                    y = matrix(avgSS, nrow = 1, ncol = numOfAPs,
+                               dimnames = list(ans$posXY, names(avgSS)))
+                    cbind(ans, y)
+                  }))
+  return(do.call("rbind", byLoc))
+}
+
+createOfflineSummary = function(df) {
+  # In order to examine distribution for all locations, angles, and are two interested APs, we
+  # will create a summary statistics for all location-orientation-AP combinations with a new factor.
+  # For each combination there are around 100 observations.
+  #
+  # Args:
+  #   df: offline data frame
+  #
+  df$posXY = paste(df$posX, df$posY, sep="-")
+
+  # create data frames for each combination
+  byLocAngleAP = with(df, by(df, list(posXY, angle, mac), function (x) x))
+
+  # summary statistic
+  signalSummary = lapply(byLocAngleAP, function(oneLoc) {
+    ans = oneLoc[1, ]
+    ans$medSignal = median(oneLoc$signal)
+    ans$avgSignal = mean(oneLoc$signal)
+    ans$num = length(oneLoc$signal)
+    ans$sdSignal = sd(oneLoc$signal)
+    ans$iqrSignal = IQR(oneLoc$signal)
+    ans
+  })
+  return(do.call("rbind", signalSummary))
+}
+
+reshapeSS = function(df,
+                     varSignal = "signal",
+                     keepVars = c("posXY", "posX", "posY"),
+                     sampleAngle = FALSE,
+                     refs = seq(0, 315, by=45)) {
+  # reshape signal strength help function
+  # aggregate signal strengths from these angles and create a data structure similar to onlineSummary
+  # Args:
+  #   df: dataframe to reshape
+  #   varSignal: variable to aggregate
+  #   keepVars: variables to retain
+  #   sampleAngle: true if we want to select one angle at random from df
+  #   refs: angle references
+  #
+  numOfAPs = length(unique(df$mac))
+  byLocation = with(df, by(df, list(posXY),
+                             function(x) {
+                               # select one angle at random for each location
+                               if (sampleAngle) {
+                                 x = x[x$angle == sample(refs, size = 1), ]}
+                               ans = x[1, keepVars]
+                               avgSS = tapply(x[, varSignal], x$mac, mean)
+                               y = matrix(avgSS, nrow = 1, ncol = numOfAPs,
+                                          dimnames = list(ans$posXY, names(avgSS)))
+                               cbind(ans, y)
+                             }))
+  return(do.call("rbind", byLocation))
+}
+
+selectTraininingData = function(newObservationAngle, df = NULL, m = 1){
+  # select observations from df to analyze aggregate signal strengths
+  # from these angles and create a data structure similar to onlineSummary.
+  #
+  # Args:
+  #   newObservationAngle: the angle of the new observation
+  #   df: offline summary data frame
+  #   m: number, between 1 and 5, of angles to keep
+
+  # m is the number of angles to keep between 1 and 5
+  refs = seq(0, by = 45, length  = 8)
+  nearestAngle = roundOrientation(newObservationAngle)
+
+  # handle odd and even m number
+  if (m %% 2 == 1)
+    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
+  else {
+    m = m + 1
+    angles = seq(-45 * (m - 1) /2, 45 * (m - 1) /2, length = m)
+    if (sign(newObservationAngle - nearestAngle) > -1)
+      angles = angles[ -1 ]
+    else
+      angles = angles[ -m ]
+  }
+  # map angles to values in refs
+  # negative angles and angles greater than 360 are mapped to appropriate angles; e.g., -45 maps to 335 and 405 maps to 45
+  angles = angles + nearestAngle
+  angles[angles < 0] = angles[ angles < 0 ] + 360
+  angles[angles > 360] = angles[ angles > 360 ] - 360
+  angles = sort(angles)
+  # select observations to analyze
+  subset = df[df$angle %in% angles, ]
+  # reshape signal strength
+  return(reshapeSS(subset, varSignal = "avgSignal"))
+}
+
+findNN = function(newSignal, trainingSubset) {
+  # want to look the distance in terms of signal strengths from these training data to the new
+  # data point. we need to calculate teh distrance from the new point to all observations in the
+  # training set with findNN().
+  # returns locations of the training observations in order of closeness to the new observation's signal strength.
+  #
+  # Args:
+  #   newSignal: signal of new observation
+  #   trainingSubset: training data to find neighbors
+  # Returns:
+  #   training neighbors
+  #
+  cols = length(trainingSubset)
+  diffs = apply(trainingSubset[ , 4:cols], 1, function(x) x - newSignal)
+  dists = apply(diffs, 2, function(x) sqrt(sum(x^2)) )
+  #closest = order(dists)
+  closest = cbind(trainingSubset[, 1:3 ], dists)
+  return(closest[order(dists),])
+  #return(trainingSubset[closest, 1:3 ])
+}
+
+predXY = function(newSignals, newAngles, training, numAngles = 1, k = 3, weighted=FALSE){
+  # Predict the XY coordinates given a list of signals along with their angles measured and training data.
+  # k neighbors will be found to make the prediction of coordinate.
+  #
+  # Args:
+  #   newSignals: list of signals to predict
+  #   newAngles: list of angles for signals to predict
+  #   trainingData: data needed for k-nn
+  #   numAngles: angles we want to use for prediction
+  #   k: number of neighbors to use for prediction
+  #   weighted: if TRUE, distances to neighbors will be used to weight XY for each neighbor to determine
+  #             location.
+  #
+  closeXY = list(length = nrow(newSignals))
+
+  for (i in 1:nrow(newSignals)) {
+    trainSS = selectTraininingData(newAngles[i], training, m = numAngles)
+    closeXY[[i]] = findNN(newSignal = as.numeric(newSignals[i, ]), trainSS)
+  }
+  # for some k of nearest neighbors, simply average the first k locations
+  # could have used weights in the average that are inversely proportional to the distance in signal strength
+  # from the test observation. will need to call findNN() to get distance to point. the weights
+  # are 1/d / sum(1/d).
+  # could also use a different metric besides Euclidean like Manhattan.
+  # could use medians instead of averages when combining neighbors if the distribution of values are quite skewed.
+  if (!weighted) {
+    estXY = lapply(closeXY, function(x) sapply(x[ , 2:3], function(x) mean(x[1:k])))
+  } else {
+    # sum X and Y weighted neighbor values.
+    estXY = lapply(closeXY, function(x) {
+      wt = (1/x[1:k, 4])/sum(1/x[1:k, 4])
+      xy = cbind(x[1:k,], wt)
+      # sum weighted neighbors
+      return(colSums(xy[1:4, 2:3] * xy[1:4, 4]))
+    })
+  }
+  return(do.call("rbind", estXY))
+}
+
+calcError = function(estXY, actualXY) {
+  # compare fit numerically with sum of squared errors
+  #
+  # Args:
+  #   estXY:
+  #   actualXY:
+  # Returns:
+  #   calculated residual sum of squares
+  return(sum(rowSums((estXY - actualXY) ^ 2)))
+}
+
+plotSSErrors = function(err, K){
+  # Plot results as sum of squared errors as function of k
+  #
+  # Args:
+  #   err: Error array
+  #   K: number of neighbors
+  #
+
+  oldPar = par(mar = c(3.1, 3.1, 1, 1), mfrow = c(1,1))
+  par(mai=c(.8, .8, .4, .25))
+
+  plot(y = err, x = (1:K),  type = "l", lwd= 2, ylim = c(900, 2100), cex.main=.8, cex.axis=.8, cex.lab=.8,
+       xlab = "Number of Neighbors",
+       ylab = "Sum of Squared Errors",
+       main = "Cross Validation Section of k")
+
+  rmseMin = min(err)
+  kMin = which(err == rmseMin)[1]
+  segments(x0 = 0, x1 = kMin, y0 = rmseMin, col = gray(0.4), lty = 2, lwd = 2)
+  segments(x0 = kMin, x1 = kMin, y0 = 800,  y1 = rmseMin, col = grey(0.4), lty = 2, lwd = 2)
+  text(x = kMin - 2, y = rmseMin + 40, label = as.character(round(rmseMin)), col = grey(0.4), cex = .6)
+
+  par(oldPar)
+}
+
+wt_signals <- function(ByAvgSig){
+
+  # Name: wt_signals
+  # Create weighted data
+  #
+  # Args:
+  #   ByAvgSig:    estimate XY line segment
+  #   
+  # Returns:
+  #   log_ang by weighted_signals matrix
+
+  # calculate signals
+  signals <- select(ByAvgSig, -posXY, -posX, -posY, -angle)
+
+  # calculate log_ang
+  loc_ang <- select(ByAvgSig, posXY, posX, posY, angle)
+
+  # calculate weighted_signals 
+  weighted_signals <- (1/signals) / rowSums(1/signals)
+
+  return(cbind(loc_ang, weighted_signals))
+}
+```

### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added table of contents and mention of dragons
I am tired.

@@
 <li>[Introduction](#Introduction)</li>
-<li>[Part 1](#Part1)</li>
+<li>[**Part 1**](#Part1)</li> 
+<ul> 
   <li>[Signal Collection](#SigCol)</li>
@@
   <li>[Discussion](#Discussion)</li>
-<li>[Part2](#Part 2)</li>
+  <li>[k-Nearest Neighbors](#knn)</li>
+</ul>
+<li>[**Part 2**](#Part 2)</li>
 
@@
 ***
+<a id='Part1'>
 
-### Part 1
+### Part 1</a>  
 #### The Approach (Explained)
@@
 To fully answer this question, we will need to explore exactly what is happening with each MAC address. The results of this exploration will define the parameters needed for such a decision.
+<a id='SigCol'>
 
-#### Here we will create a Signal Collection Map
+#### Here we will create a Signal Collection Map</a>
 For this exercise, we will plot both the count and locations of the signals recorded for the access points.
@@
 ***
+<a id='SigStren'>
 
-#### Signal Strength Distributions
+#### Signal Strength Distributions</a>
 Plotting the individual signal strength distributions for each AP angle will help add perspective to the information we are working with.
@@
 ***
+<a id='SigDen'>
 
-#### Time to investigate Signal Density
+#### Time to investigate Signal Density</a>
 To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each 45 degree angle.
@@
 ***
+<a id='Discussion'>
 
+### Discussion</a>
 #### We now have enough information to address these questions before an analysis of kNN is conducted:
@@
 
-The evidence would suggest both addresses should be kept within the analysis to increase the coverage for the entire spectrum of strengths.
-The address chosen to be rejected not only has a lower signal strength when measured from either end of the building, it also shows to have a higher signal density when it is further away from the receiver.
-The address chosen to be kept has a higher signal density closer to the receivers.
+The evidence thus far would suggest both addresses should be kept within the analysis to increase the coverage for the entire spectrum of strengths.
+The address chosen to be rejected not only has a lower signal strength when measured from either end of the building, but it also shows to have a higher signal density when it is further away from the receiver.
+The address chosen to be kept has a higher signal density closer to the receivers with the accompanying higher signal strengths.
 By keeping both addresses, the model should be able to generate a more accurate prediction of the layout of the building.
-This theory will be tested as we tour through k-Nearest Neighbors.
+This theory will be tested as we tour k-Nearest Neighbors.
 
-#### Let's dive into the K-Nearest Neighbors:
+***
+
+<a id='knn'>
+
+#### Let's dive into the K-Nearest Neighbors:</a>
+
 <font color = "red"><b>Discuss KNN here</b></font>
@@
 
+With the unique MAC addresses in hand, an accurate profile of the different signal strengths reveived by angle is able to be determined. This can be done with the following code: 
 ```{r}
@@
 This output shows the signal strengths at each angle. This will be useful later!
-<font color = "red"><b>&uarr;Explain why &uarr;</b></font>
+<font color = "red"><b>&larr;Explain why &larr;</b></font>
 
@@
 
+<font color = "purple">
+(this can probably be removed)
+
 > for k-NN, we will include training data with angles close to point in question since angle matters.
@@
 > instead return a set of mx166 signals for each access point?
+</font>
 
@@
 
+<font color="red"><b>&uarr;We will need to discuss these results&uarr;</b></font>
+
 <font color = "green" size = "5px"><b>This is where we will perform cross validation and plot for desired AP for a range of k's </b></font>
@@
 ```
+
+### **Here be Dragons &darr;**
 
@@
 
-#Part 2: Alternatvie k-Nearest Neighbors
+#Part 2: Alternative k-Nearest Neighbors
 </a>

### 2017-06-19 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added section discussing power-based location prediction.

@@
 
-### Table of Contents:
+### Table of Contents: {#TOC}
 <li>[Assignment Requirements](#AssignmentOutline)</li>
@@
     + *Use calcError() to compare approaches*
+
+[&uarr;TOC](#TOC)
 
@@
 </a>
+
+##2.1: Examination of Power-based search
+The case study document mentioned the curvature in the scatterplots of signal strength versus distance, and suggested using a log function to attempt to remove the curve.
+Since the signal strength data are reported in dBm, they are already the result of a log function: $dBm = 10*log_{10}(power/1 mW)$; therefore, we converted the reported signal strength values back to a power value and examined the effect that had on location prediction.
+Many of the functions used had to be modified to include the power conversion: $power = 10^{(signal/10)}$ and to use the new field in the calculations.
+Examination of surface plots of power show that the power value tends to isolate the AP more definitively than signal strength because $power \propto 1/distance^2$.
+
+![Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
+![Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)
+
+The thought was that by having such a dramatic drop in power over a shorter range, we should be able to reduce the variability in predicted location and get better results.
+However, when we simply substituted power for signal strength in the existing functions, the predictions using power were worse than using signal strength.
+![Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
+
+Using different values for 'k', we had these estimated errors:
+
+| k |Mean distance error|Book-reported error|
+|:-:|:-----------------:|:-----------------:|
+| 1 |779.8603|659|
+| 3 |598.8870|307|
+| 4 |546.8603||
+| 5 |640.8003|276|
+
+Even though the power-based method better isolated the individual APs, we believe that the search algorithm should be redesigned to use all APs, instead of just one or two, due to the power drop-off.
+As the distance from the reference AP increases, the power drops off so much that it becomes difficult to narrow down the actual location.
+By redesigning the algorithm to use all APs, we may be able to capitalize on the power drop-off to better hone in on the target's location.
+
+##2.2: Weighted K-Nearest Neighbors
+
 ```{r}

### 2017-06-19 Section2_TE.rmd
Created a rmd file for section 2

##2.1: Examination of Power-based search
The case study document mentioned the curvature in the scatterplots of signal strength versus distance, and suggested using a log function to attempt to remove the curve. Since the signal strength data are reported in dBm, they are already the result of a log function: $dBm = 10*log_{10}(power/1 mW)$; therefore, we converted the reported signal strength values back to a power value and examined the effect that had on location prediction.
Many of the functions used had to be modified to include the power conversion: $power = 10^{(signal/10)}$ and to use the new field in the calculations.
Examination of surface plots of power show that the power value tends to isolate the AP more definitively than signal strength because $power \propto 1/distance^2$.

![Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
![Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)

The thought was that by having such a dramatic drop in power over a shorter range, we should be able to reduce the variability in predicted location and get better results.  However, when we simply substituted power for signal strength in the existing functions, the predictions using power were worse than using signal strength.

![Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)

Using different values for 'k', we had these estimated errors:

| k |Mean distance error|Book-reported error|
|:-:|:-----------------:|:-----------------:|
| 1 |779.8603|659|
| 3 |598.8870|307|
| 4 |546.8603||
| 5 |640.8003|276|

Even though the power-based method better isolated the individual APs, we believe that the search algorithm should be redesigned to use all APs, instead of just one or two, due to the power drop-off.  As the distance from the reference AP increases, the power drops off so much that it becomes difficult to narrow down the actual location.  By redesigning the algorithm to use all APs, we may be able to capitalize on the power drop-off to better hone in on the target's location.

##2.2: Investigating Weights

The description of the weighting procedure brought up some questions about the meaningfulness of the approach.  This led to a side investigation of how the nearest-neighbors method might be affected by weights.

The investigation involved exporting the offine summary dataset and the online dataset to CSV files for use in Excel.  Then we looked at the process manually - taking the first report in the online dataset that matched our reference AP and looking for nearest neighbors in the offline summary dataset.  The target record stated the true position to be 0, 0.05 (lower left corner of the building).  The reference AP is within line of sight of the target device at coordinates 1,14.

First, we selected only those records in the offline summary that matched the AP, then further filtered to those records that matched orientation.  Plotting the measured signal strength for the AP and orientation, we can see that there is a fairly extensive swath of positions in which the target signal strength can occur.

![*Signal strength from AP @ 1,14*](SignalStrength1.png).

Using the Excel function SMALL, we can find the n smallest differences - these would be the nearest neighbors.

|Record#|Sig Str Diff|Pos X|Pos Y|
|:-----:|:----------:|:---:|:---:|
| 3767  | 0.0196078  | 4   | 7   |
|  357  | 0.0202128  | 1   | 2   |
|  117  | 0.3846154  | 2   | 0   |
| 3817  | 0.9354839  | 3   | 7   |
|  505  | 1.0510204  | 1   | 3   |

Avg X = 2.2; Avg Y = 3.8; True X = 0.0; True Y = 0.05; Error = 4.3477

We can see that the target's reported signal strength occurs throughout the corridor as shown in the signal strength plot above.  Averaging those values pulls the estimate away from the true area.

If we consider weighting the neighbors based on physical distance (as was suggested in the case study), the points closer to the AP will be weighted higher than the points further away, which would (in this particular example) pull the estimated location further away from the true location and closer to the AP, which would increase the error.  This is what led to this investigation in the first place.

The case study also hints at using multiple APs to do the neighbor selection.  If, for example, we chose the AP at the opposite corner of the building (the lower right corner), and examined the same target report as before, we can see an additional problem:

![*Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)

Now if we look for just simple differences in signal strength, we can get points for either AP that could be spread across the entire building.

|Record#|Sig Str Diff|Pos X|Pos Y|AP|
|:-----:|:----------:|:---:|:---:|:-:|
|  670  | 0.0625     | 4   | 3   |1, 14|
|  2568 | 0.0919540  | 15  | 8   |33.5, 2.8|
|  856  | 0.1034483  | 11  | 4   |33.5, 2.8|
|  936  | 0.1263158  | 12  | 5   |33.5, 2.8|
|  408  | 0.1612903  | 0   | 7   |1, 14|

Avg X = 8.4; Avg Y = 5.4; True X = 0.0; True Y = 0.05; Error = 9.959

Again, weighting the neighbors on physical distance reinforces a location closer to the AP at 33.5, 2.8, which increases the error from the true location.

One possible approach is to use the simultaneous measurements as a group and find nearest neighbors that exhibit similar patterns.  For example the test record we have used here is part of a series of measurements taken at the same time and orientation.  The highest average signal strength was against AP 97:8A (-43.25 dBm) and the lowest average signal strength was against AP 97:8D (-63 dBm). If we search for that pattern in the training data, we get a more reasonable set of neighbors:

|Timestamp|97:8A signal|97:8D signal|Signal "Dist"|Pos X|Pos Y|
|:-------:|:----------:|:----------:|:-----------:|:---:|:---:|
|02:45:43 | -43 | -63 | 0.25  |1 | 5 |
|02:14:34 | -44 | -63 | 0.75  |1 | 4 |
|01:06:28 | -44 | -61 | 2.136 |0 | 1 |
|06:31:15 | -41 | -63 | 2.25  |2 | 12|
|00:35:27 | -41 | -64 | 2.462 |0 | 0 |

Avg X = 0.8; Avg Y = 4.4; True X = 0.0; True Y = 0.05; Error = 4.423

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
revised intro

@@
 ### <b>Assignment</b> </a>
-1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both the chosen and rejected Access Points:
+1. Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both the *chosen* and *rejected* Access Points:
     + *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
-    + *Which MAC address yields the best prediction of location? (Including determining locations by using data corresponding to both MAC addresses)*
+    + *Which MAC address yields the best prediction of location? (Including determining locations by using data corresponding to both MAC addresses)?*
     + *Does using data for both MAC addresses simultaneously yield more, or less accurate prediction of location?*
 
-2. Implement alternative k-nearest precition method using weights on received signal strength.
+2. Implement alternative k-nearest prediction method using weights on received signal strength.
     + *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
-    + *Use calcError() to compare approaches*
+    + *Use calcError() to compare approaches.*
 
@@
 
-For this case study, the overall goal is to use K-Nearest Neighbors (kNN) and an alternative weighted method to be able to determine how different handheld devices moved through a building <font color = "red">(at Dartmouth.
-Let's confirm this and add some more detail)</font>.
-This can be achieved with an analysis of how the different Media Access Control (MAC) addresses for each of the devices interacted with the Access Points (APs) of the building.
-In this case, the APs are routers with static locations throughout the building.
+The case study in chapter 1 of the book Data Science in R by Nolan and Lang presented an indoor posiitoning system (IPS) for WiFi devices on a floor of a building at the University of Mannheim.
+The IPS was trained and tested with data collected at the University with a hand-held device on a grid of 166 points spaced one meter apart in the hallways.
+The algorithm for the IPS is the K-Nearest Neighbors (k-NN).
+
+The original investigators used a hand-held scanning device at each point and recorded a timestamp, mac id of the device, position of the device, the orientation of the device, and the following data for each WiFi access point (AP): recevied signal strength, channel frequency, and device mode of operation.
+The WiFi APs were located throughout the building and identified by their unique Media Access Control (MAC) address.
+
+<font color = "red"><b>
+- Need example training input line
+- Need table of training data: name, description, type, example
+</b></font>
 
@@
 
-The access points as stated above allow for the tracking of the devices thanks to the logged history of times, signal strengths and angles for each MAC address with respect to each access point.<font color = "red">&larr;(consider fragmenting this sentence)</font> From the information that has been given, there are known to be six APs on the building floor in question.
-However, this is a problem since there are seven APs recorded.
-To determine which AP to exclude from the analysis, we will need to compare these two addresses:
-
-    chosenAP   = "00:0f:a3:39:e1:c0"
-    rejectedAP = "00:0f:a3:39:dd:cd"
-
-and decide which to exclude and why.
-
-<font color = "purple"><b>
-- Discuss why we're trying to remove one
-</b></font>
-
-Removing one of these APs will allow us to remove noise from the resulting model.
-Since only six APs are recorded as existing on the floor, the presence of a seventh points to external interference.
-Once the superfluous address has been identified, we will execute testing to determine whether there will be any benefit added to the model by keeping or excluding said data.
-This decision will ultimately lead to a better prediction of location of each device within this floorplan.
+Six APs were initially discovered in the original anlysis in the case study.
+The authors then decided to reject AP "00:0f:a3:39:dd:cd" (rejectedAP) stating it was identified as an extra address.
+The details of their reasoning was not clearly explained.
+The authors chose AP "00:0f:a3:39:e1:c0" (chosenAP) instead.
+This case study will evalute the author's decision and decide if the IPS system would be more accurate with the chosenAP, rejectedAP, or both APs.
 
@@
 
-K-Nearest Neighbors (kNN) is an ideal method to use in this case as the goal is to be able to predict the path of each device on their journey through the building.
-To utilize this method, we use separate testing and training data.
-The training data are known observations from well documented locations of both the handheld devices and APs.
-These locations are represented by the angle at which the AP received the signal from the device as well as the overall signal strength received.
+k-NN algorithm is an ideal method because we have an abundant training data and our prediction is not based on parametric relationship with the data.
+It is based on proximity.
+The downside is the model must hold the entire training data set for maximum predictability.
+The k-NN model is trained on the training data and then tested on the test data to determine its accuracy.
+The number of neighbors to used for location prediction is adjusted to reduce the accuracy.
 
-With this training data, each observation is evaluated along with the observations which occur in the general vicinity in the training data.
-These occurrences allow for the different probabilities of occurrence to be proposed through the model.
-Once this model has been generated for k neighbors, it is then tested against the unobserved test data and adapted to suit what actually occurs.
-This allows for an accurate gauge of accuracy of the model and is recreated for a span of possible k neighbors to exhaust the total possible number of observations.
-
-<font color = "purple"><b>
-- Discuss the alternative weighted method. 
+<font color = "red"><b>
+- can we still a graphic from scikit-learn of k-NN model?
 </b></font>
 
-In contrast to the kNN method, another option to explore is a weighted method based on the signal strength received at the access point.
-This method involves defining a new metric where the weight of each observation is inversely proportional to the distance from the test observation.
-From a topographical standpoint, the observations which are closest to the observation will have a much higher influence than those which exist further away.
-
-<font color = "purple"><b>
-- Discuss potential for prediction and what we expect to come of this.
-</b></font>
-
-The evaluation of this data will allow us to determine which data will be critical to the creation of an accurate model and which will not.
-Since we have six access points and seven addresses, this thorough investigation will help us decide which address is simply an external interference and if it is beneficial to keep involving said data.
-As a result, we expect to have a more thorough understanding of the difference between k-Nearest Neighbors and the weighted alternative.
+In part 1 of the case study, the author's original decision to rejected an AP is revisited to see if this was the best decision.
+In part 2, the prediction of X,Y based on the average of its neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions.
+Alternate weights will also be analyzed.
 

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added code to 2.2

@@
 
+Prediction based on weighted X and Y &darr;
+
 ```{r}
-# create weights for neighbors based on signal strength and multiple times X and Y and sum neighbors
-# to get X and Y.
 estXY = predXY(newSignals = onlineSummary[ , 6:12],
-               newAngles = onlineSummary[ , 4],
-               offlineSummary, numAngles = 3, k = 5, weighted=TRUE)
-actXY = onlineSummary[ , c("posX", "posY")]
+               newAngles = onlineSummary[, 4],
+               offlineSummary, numAngles = 3, k = 5, weighted = TRUE, exp = .5)
+actXY = onlineSummary[, c("posX","posY")]
+calcError(estXY, actXY)
+```
+
+&darr; this is equivalent to mean SS error result of 210
+
+```{r}
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
+               newAngles = onlineSummary[, 4],
+               offlineSummary, numAngles = 3, k = 5, weighted = TRUE, exp = .001)
+actXY = onlineSummary[, c("posX","posY")]
 calcError(estXY, actXY)
@@
 ***
+***
+
 <a id='References'>
@@
 This is the code used to define the functions used within the study. It can be found under 'MSDS7333-baldree-case6-fx.r' which is included in the final submission  
-```{r}
+```{r eval = FALSE, echo=FALSE}
 processLine = function(x)

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly_BBrock-Update.Rmd
add text for weighted K-NN
add text for weighted K-NN

--- 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
@@
 ##2.2: Weighted K-Nearest Neighbors
+
+In this exercise, we investigated using an alternative k-nearest prediction method using the weighted 
+K-nearest neighbors approach.
+The calculation of the Euclidean distance error was found to be 1194, whereas the Mean SS error was found to be 210.
+The Euclidean distance weighted k-nearest neighbors
+method had the highest error, while the Mean SS weighted K-nearest neighbors method had the smaller
+error.
+
+| k |Euclidean distance error|Mean SS error|
+|:-:|:-----------------:|:-----------------:|
+| |1194|210|
 

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
update

--- 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
@@
 The IPS was trained and tested with data collected at the University with a hand-held device on a grid of 166 points spaced one meter apart in the hallways.
-The algorithm for the IPS is the K-Nearest Neighbors (k-NN).
+The algorithm for the IPS is the K-Nearest Neighbors (k-NN) [1].
 
@@
 In part 2, the prediction of X,Y based on the average of its neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions.
-Alternate weights will also be analyzed.
+Alternate weights will alsobe analyzed.
 
 ***
-<a id='Part1'>
+# <a id='Part1'>Part 1: Access Point Analysis</a>  
+Majority of the code in this case study was leveraged from [1]. The code was refactored for readibility and modified for part 1 and 2 analysis.
 
-### Part 1</a>  
-#### The Approach (Explained)
+## 1.1 Read and Clean the Data
 
-Before we begin, we will need to set the parameters for printing and formatting
+Set options parameters for printing and formatting.
 ```{r}
 options(digits = 2)
-options(error = recover, warn = 2)
+options(error = recover, warn = 1)
 ```
 
-Before continuing, we will need to address another file which houses the predefined functions. This helps to save notebook real estate and prevents confusion:
+Include helper functions defined in another notebook.
 ```{r}
@@
 
-Read offline data
+Read in offline data used for training our model.
 ```{r}
@@
 
-Conduct a more thorough data analysis into these two MAC addresses, including determining locations by using data corresponding to both:
+## 1.2 Analyze the Two Acess Points
 
-    * chosenAP = "00:0f:a3:39:e1:c0"
-    * rejectedAP = "00:0f:a3:39:dd:cd"
-
-Since we have the two MAC addresses in question, we will declare these addresses as variables to simplify later code.
+Define the access point (AP) mac addresses that will be analyzed. We also define a variable for all remaining APs.
 
@@
 
-#### * *Which of these two MAC addresses should be used and which should not be used for Real Time Location System (RTLS)?*
+### 1.2.1 <a id='SigCol'>Signal Map</a>
+Plot the total signals captured for each AP to determineif the coverage was similar. A severe gap in the number of points captured could be a reason for lower predictability.
 
-To fully answer this question, we will need to explore exactly what is happening with each MAC address. The results of this exploration will define the parameters needed for such a decision.
-<a id='SigCol'>
-
-#### Here we will create a Signal Collection Map</a>
-For this exercise, we will plot both the count and locations of the signals recorded for the access points.
-Even though it would be ideal to plot the access points(APs) directly on the map, we do not know their X and Y coordinates.
+The total signals for each AP mac in question, chosenAP and rejectedAP, are plotted at their respective X,Y point.
+We are unable to plot the access points (APs) directly on the map because we do not know their X, Y coordinates.
 
@@
 
-At first glance, we can see that the maps are quite similar.
-Unfortunately, the information gained from a visual comparison of these two diagrams does not give us a significant insight into which MAC address should be kept and which should be rejected.
-The next step to reach our goal will be to look at the individual Signal Strength Distributions
+The maps show there were near equal amount of training points captured for both APs. A visual comparison of 
+these two diagrams does not give us a significant insight into which MAC address should be kept and which should be rejected. Now, let us examine the distribution strength of the APs.
 
-***
-<a id='SigStren'>
-
-#### Signal Strength Distributions</a>
-Plotting the individual signal strength distributions for each AP angle will help add perspective to the information we are working with.
-For the sake of consistency, we will use the same stationary coordinates listed in the course book (coord : (2, 12)) <font color = "red">(Add reference here.)</font>
+### 1.2.1 <a id='SigStren'> Signal Strength Distributions</a>
+Plot the signal strength distributions for each AP per angle for better understanding at location (X=2, Y=12).
+A more thorough analysis would examine the distributions at all locations points, but time did not permit this in depth analysis.
 
@@
 
-From the literature, we know the higher the decibel-milliwatts means a stronger signal strength.
-A stronger signal strength indicates a closer physical position to the AP since signal strengths decrease exponentially with distance.<font color="red">Find this in the book and cite it.</font> Comparing the signal strength distributions in this chart, we see "e1:c0" has a stronger signal strength at this location than "dd:cd."
+The higher the decibel-milliwatts means a stronger signal strength which indicates a closer physical position to the AP since signal strengths decrease exponentially with distance.<font color="red">Find this in the book and cite it.</font> The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage.
+What would this plot look like if we moved the fixed location point on the other side of the building at (X=33, Y=3)?
 
-What if we compared the two addresses from the opposite side of the building? 
 ```{r}
@@
 ```
-Through the use of another perspective, the MAC address with the weaker signal at (2,12) still has a relatively weak signal when viewed from an AP at the other side of the building.
-The evidence against the address with the lower signal strength is mounting.
+This plot is similar to the previous distribution but with no outliers.
+At this point we know signal strength is different for the two access points.
+We don't know if this will effect our prediction accuracy.
+Now, let us analyzed the two APs signal strength distribution over angles.
 
-***
-<a id='SigDen'>
-
-#### Time to investigate Signal Density</a>
-To properly investigate the difference between these two addresses, the signal strength densities for each address should be plotted for each 45 degree angle.
-This will help display the defining characteristics for each address.
-The location used for this analysis is the (24,4) from the book.
+### 1.2.2 <a id='SigDen'> Signal Density</a>
+Plot signal strength frequency per capture angle for normalcy.
 
@@
 ```
-From this plot, it can be seen that the rejectedAP has a higher signal density at lower strength while the acceptedAP has a higher density coverage at a higher strength.
-Judging by the distributions themselves, it seems as if many of the plots are normally distributed while some of the other plots (i.e.
-"dd:cd" at 180 and "e1:c0" at 315) are heavily skewed or even bimodal.
+The above plot, shows the rejectedAP "dd:cd" has a higher signal density at lower strength while the acceptedAP "e1:c0" has a higher density coverage at a higher strength.
+The distributions show that many of the plots are normally distributed while some of them; i.e., "dd:cd" at 180 and "e1:c0" at 315 are heavily skewed or even bimodal.
 Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle.
 This would suggest the median of each distribution varies by angle and attributes to a difference in perceived strength.
-
-***
+This means the angle of the device contributes to the APs strength of signal.
 
 #### Average Signal Strength Distributions
-To examine the distribution for all locations and angles of the two APs in question, we will generate the summary statistics.
-The summary statistics for all possible combinations of locations and orientations for each AP through the use of a new factor.
+Plot the signal strengh distribution for all locations through the average signal strenth for all locations by angle. Summary statistics for all possible locations by angle was created.
 
@@
 
-From this figure, it is apparent that there is considerable overlap between the two distributions of signal strength.
-This is evidence in favor of keeping both addresses in the analysis to provide greater coverage.
+This above figure shows that there is considerable overlap between the two distributions of signal strength. This suggest that there may be value in keeping both APs in the k-NN model.
 
-***
-<a id='Discussion'>
+### 1.2.3 <a id='Discussion'>Discussion</a>
+The data analysis shows that both APs are very similar except for the signal strength coverage.
+We do not have enough information at this time to determine if one AP is better or worse than the other.
+Therefore, we will examine each one in the k-NN model along with both of them to determine which option yields the best accuracy.
 
-### Discussion</a>
-#### We now have enough information to address these questions before an analysis of kNN is conducted:
+### 1.2.4 <a id='knn'>K-Nearest Neighbors</a>
+k-NN is a non-parametric machine learning algorithm that is instance-based learning as opposed to generalize learning.
+Our first implementation of k-NN learns based on teh number of neighbors within signal strength distance to the point in question from the training points.[2]
 
-* **Which MAC address yields the best prediction of location? (Include determining locations by using data corresponding to both MAC addresses)**
-* **Does using data for both MAC addresses simultaneously yield a more or less accurate prediction of location?**
-
-The evidence thus far would suggest both addresses should be kept within the analysis to increase the coverage for the entire spectrum of strengths.
-The address chosen to be rejected not only has a lower signal strength when measured from either end of the building, but it also shows to have a higher signal density when it is further away from the receiver.
-The address chosen to be kept has a higher signal density closer to the receivers with the accompanying higher signal strengths.
-By keeping both addresses, the model should be able to generate a more accurate prediction of the layout of the building.
-This theory will be tested as we tour k-Nearest Neighbors.
-
-***
-
-<a id='knn'>
-
-#### Let's dive into the K-Nearest Neighbors:</a>
-
-<font color = "red"><b>Discuss KNN here</b></font>
-
-First and foremost, let's display the unique MAC addresses: 
+Print out the unique AP MAC addresses showing that we have six of them.
 ```{r}
@@
 
-With the unique MAC addresses in hand, an accurate profile of the different signal strengths reveived by angle is able to be determined. This can be done with the following code: 
+Table the test data to review the frequency for each angle and each coordinate. 
 ```{r}
@@
 ```
-This output shows the signal strengths at each angle. This will be useful later!
-<font color = "red"><b>&larr;Explain why &larr;</b></font>
+The above output shows that each angle has test data meaning we can test all angle for accuracy.
 
-Organize the data where each AP is in a column. 
+Reshape the test data where each AP is in a column. 
 ```{r}
@@
 
-<font color = "purple">
-(this can probably be removed)
-
-> for k-NN, we will include training data with angles close to point in question since angle matters.
-> if we want one angle, then include angles that match the rounded orientation of new observation.
-> if we want two angles, then pick two multiples of 45 degrees that flank the new observation's orientation.
-> if we want three angles, then pick the closest 45 degree increment and on either side of it.
-> what if we didn't want to collapse the signal strengths across the m angles, and
-> instead return a set of mx166 signals for each access point?
-</font>
-
-When it comes to the kNN, we will need to include the data which include the angles close to the point in question.
-As observed before, the angle matters!
-For this exercise, we will need to abide by the following outlines:
+When it comes to the kNN, we need to include the data which include the angles close to the point in question.
+As observed before, angle matters.
+For this exercise, we need to abide by the following rules:
 
@@
 
-What if we didn't want to collapse the signal strengths across the *m* angles and, instead, we return a set of mx166 signals for each access point?
-
-Here, we begin by iterating through 1-3 neighbors at 1-3 angles for each combination to determine which will be best suited for cross-validation:
+Now, we analyze or three scenarios for both APs included, rejectedAP left out, and then chosenAP left out.
+Three neighbors were used for three angles.
+The accuracy for each scenario as determined by the sum of squares difference between estimated XY and actual XY were outputted as well.
+Cross-validation was not used at this time.
 
@@
 ```
+The above results show that including both APs yielded a better prediction accuracy followed by keeping chosenAP and then rejectedAP.
+The author's choice to keep chosenAP was indeed better than rejectedAP, but keeping both of them is even better.
 
-<font color="red"><b>&uarr;We will need to discuss these results&uarr;</b></font>
+<font color = "Red">talk about cross-validation</font>
 
-<font color = "green" size = "5px"><b>This is where we will perform cross validation and plot for desired AP for a range of k's </b></font>
+We will now perform cross-validation with both APs to determine the best k.
+Three angles will be used.
+Future work would include parameterization of the angles from 1 to 3 to determine to determine the best m.
 
-Perform cross validation and plot for desired AP for a range of k's for each fold. We want to find k-NN estimates from 1 to k and aggregate errors over the v folds.
-
-Choose a v-fold of 11
+#### Cross-Validation
+Choose a v-fold of 11.
 ```{r}
@@
 ```
-
-### **Here be Dragons &darr;**
 
@@
 ```
-<font color = "Red"><b>&uarr; come back to this one &uarr; For some reason it only runs after it throws an error (e.g.change matrix to sapply, run, change back and run again</b></font>
 
@@
 ```
+Given the plot above, the best k shows 9, but this changes if you rerun the cross-validation. Experimentation with k yielded 5 with the lowest error of 210 so it was used below.
 
@@
 
-*** 
-<a id='Part2'>
+<font color = "Red">provide some feel good about part 1.
+we concluded that keeping both APs yields better accuracy with a k-NN of 5 with 3 angles.
+We did not iterate of the angles to determine best parameter (arg!).
+The training data neighbors to the test data was determined by the signal strength distance and then the neighbors average X,Y was used to predict the test point location.
+</font>
 
-#Part 2: Alternative k-Nearest Neighbors
-</a>
+# <a id='Part2'> Part 2: Alternative k-Nearest Neighbors</a>
 
-##2.1: Examination of Power-based search
+## 2.1: Examination of Power-based search
 The case study document mentioned the curvature in the scatterplots of signal strength versus distance, and suggested using a log function to attempt to remove the curve.
@@
 
-##2.2: Weighted K-Nearest Neighbors
+## 2.2: Weighted K-Nearest Neighbors
 
@@
 ***
-<a id='References'>
+# <a id='References'>References</a>
+[1] Nolan, Deborah; Data Science in R; 2015; Predicting Location via Indoor Positioning Systems; Chapter 1. 
+[2] Scikit-learn k-NN; http://scikit-learn.org/stable/modules/neighbors.html#nearest-neighbor-algorithms.
 
-### References  
-</a>
-
-<a id='Appendix'>
-
-### Appendix (including function code)  
-</a>
-This is the code used to define the functions used within the study. It can be found under 'MSDS7333-baldree-case6-fx.r' which is included in the final submission  
-```{r}
+# <a id='Appendix'>Appendix (including function code)</a>
+This is the code used to define the functions used within the study. It can be found under 'MSDS7333-baldree-case6-fx.r' which is included in the final submission. 
+```{r eval=FALSE, echo=FALSE}
 processLine = function(x)

### 2017-06-20 MSDS7333-baldree-case6.r

@@ -140,7 +140,7 @@ for (ap in c("None", rejectedAP, chosenAP)){
 }
 
 # perform cross validation and plot for desired AP for a range of k's
-# for each fold, we want to find k-NN estimates from 1..K and aggregate errros over the v folds.
+# for each fold, we want to find k-NN estimates from 1..K and aggregate errors over the v folds.
 # choose a v-fold of 11
 v = 11
 # number of locations

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Incorporated section 2 materials, grammar and spelling checks, section/figure/table numbering

@@
 </ul>
-<li>[**Part 2**](#Part 2)</li>
-
+<li>[**Part 2**](#Part2)</li>
+<ul> 
+  <li>[2.1 Examining Power](#ExamPower)</li>
+  <li>[2.2 Examining Weights](#ExamWeight)</li>
+  <li>[2.3 Weighted kNN](#WtKNN)</li>
+</ul>
 <li>[References](#References)</li>
@@
 
-2. Implement alternative k-nearest prediction method using weights on received signal strength.
+2. Implement alternative k-nearest neighbors prediction method using weights on received signal strength.
     + *For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?*
@@
 
-The case study in chapter 1 of the book Data Science in R by Nolan and Lang presented an indoor posiitoning system (IPS) for WiFi devices on a floor of a building at the University of Mannheim.
+The case study in chapter 1 of the book *Data Science in R* by Nolan and Lang presented an indoor posiitoning system (IPS) for WiFi devices on a floor of a building at the University of Mannheim.
 The IPS was trained and tested with data collected at the University with a hand-held device on a grid of 166 points spaced one meter apart in the hallways.
-The algorithm for the IPS is the K-Nearest Neighbors (k-NN) [1].
+The algorithm for the IPS is the K-Nearest Neighbors (k-NN) [1](#REF01).
 
-The original investigators used a hand-held scanning device at each point and recorded a timestamp, mac id of the device, position of the device, the orientation of the device, and the following data for each WiFi access point (AP): recevied signal strength, channel frequency, and device mode of operation.
+The original investigators used a hand-held scanning device at each point and recorded a timestamp, MAC id of the device, position of the device, the orientation of the device, and the following data for each WiFi access point (AP): recevied signal strength, channel frequency, and device mode of operation.
 The WiFi APs were located throughout the building and identified by their unique Media Access Control (MAC) address.
@@
 - Need example training input line
+
+t=1139643118358;id=00:02:2D:21:0F:33;pos=0.0,0.0,0.0;degree=0.0;00:14:bf:b1:97:8a=-38,2437000000,3;00:14:bf:b1:97:90=-56,2427000000,3;00:0f:a3:39:e1:c0=-53,2462000000,3;00:14:bf:b1:97:8d=-65,2442000000,3;00:14:bf:b1:97:81=-65,2422000000,3;00:14:bf:3b:c7:c6=-66,2432000000,3;00:0f:a3:39:dd:cd=-75,2412000000,3;00:0f:a3:39:e0:4b=-78,2462000000,3;00:0f:a3:39:e2:10=-87,2437000000,3;02:64:fb:68:52:e6=-88,2447000000,1;02:00:42:55:31:00=-84,2457000000,1
+
 - Need table of training data: name, description, type, example
+
+|Data Element|Description|Type|Example|
+|:-|:-|:-|:-|
+|t |timestamp (milliseconds since midnight 1/1/1970)|long integer|t=1139643118358 --> 2/11/2006 07:35:28|
+|id|target device MAC address|6-octet sequence in hexadecimal|id=00:02:2D:21:0F:33|
+|pos|coordinates of the target device in meters from a choke point in the host building|3 floating point values, comma-separated|pos=0.0,0.0,0.0|
+|degree|Orientation of the target device (with respect to North) in degrees|float|degree=0.0|
+|MAC Address|one or more reports from various Access Points|comma-separated value list|00:14:bf:b1:97:8a=-38,2437000000,3|
+
+|Data Element|Description|Type|Example|
+|-|-|-|-|
+|signal|The signal strength reported for the Access Point in dBm|integer|-38|
+|channel|The frequency of the channel used by the Access Point in Hz|integer|2437000000|
+|mode|The mode of operation for the Access Point (1 = Ad Hoc, 3 = Access Point)|integer|3|
+
 </b></font>
@@
 
-Six APs were initially discovered in the original anlysis in the case study.
+Six APs were initially discovered in the original analysis in the case study.
 The authors then decided to reject AP "00:0f:a3:39:dd:cd" (rejectedAP) stating it was identified as an extra address.
@@
 
-k-NN algorithm is an ideal method because we have an abundant training data and our prediction is not based on parametric relationship with the data.
+The k-NN algorithm is an ideal method because we have an abundant training data set and our prediction is not based on parametric relationship with the data.
 It is based on proximity.
-The downside is the model must hold the entire training data set for maximum predictability.
+The downside is the model must hold the entire training data set in memory for maximum predictability.
 The k-NN model is trained on the training data and then tested on the test data to determine its accuracy.
-The number of neighbors to used for location prediction is adjusted to reduce the accuracy.
+The number of neighbors to be used for location prediction is adjusted to improve the accuracy.
 
@@
 
-In part 1 of the case study, the author's original decision to rejected an AP is revisited to see if this was the best decision.
+In part 1 of the case study, the author's original decision to reject an AP is revisited to see if this was the best decision.
 In part 2, the prediction of X,Y based on the average of its neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions.
-Alternate weights will alsobe analyzed.
+Alternate weights will also be analyzed.
+
+[&uarr;TOC](#TOC)
 
@@
 # <a id='Part1'>Part 1: Access Point Analysis</a>
-Majority of the code in this case study was leveraged from [1]. The code was refactored for readibility and modified for part 1 and 2 analysis.
+
+The majority of the code in this case study was leveraged from [1](#Ref01). The code was refactored for readibility and modified for part 1 and 2 analysis.
 
@@
 
-Define the access point (AP) mac addresses that will be analyzed. We also define a variable for all remaining APs.
+Define the access point (AP) MAC addresses that will be analyzed. We also define a variable for all remaining APs.
 
@@
 ```
+[&uarr;TOC](#TOC)
 
 ### 1.2.1 <a id='SigCol'>Signal Map</a>
-Plot the total signals captured for each AP to determineif the coverage was similar. A severe gap in the number of points captured could be a reason for lower predictability.
+Plot the total signals captured for each AP to determine if the coverage was similar. A severe gap in the number of points captured could be a reason for lower predictability.
 
-The total signals for each AP mac in question, chosenAP and rejectedAP, are plotted at their respective X,Y point.
+The total signals for each AP MAC in question, chosenAP and rejectedAP, are plotted at their respective X,Y point.
 We are unable to plot the access points (APs) directly on the map because we do not know their X, Y coordinates.
@@
 plotSignalMaps(c(chosenAP, rejectedAP))
+#  figure 1 - number of reports at each location for the rejected and chosen APs.
 ```
 
-The maps show there were near equal amount of training points captured for both APs. A visual comparison of 
-these two diagrams does not give us a significant insight into which MAC address should be kept and which should be rejected. Now, let us examine the distribution strength of the APs.
+The maps show there were near equal amounts of training points captured for both APs. A visual comparison of 
+these two diagrams does not give us sufficient insight into which MAC address should be kept and which should be rejected. Now, let us examine the distribution strength of the APs.
+
+[&uarr;TOC](#TOC)
 
@@
 Plot the signal strength distributions for each AP per angle for better understanding at location (X=2, Y=12).
-A more thorough analysis would examine the distributions at all locations points, but time did not permit this in depth analysis.
+A more thorough analysis would examine the distributions at all locations points, but time did not permit this in-depth analysis.
 
@@
 plotBoxplotSignalStrength(df)
+# figure 2 - Signal strength distribution across both Access Points
 ```
 
-The higher the decibel-milliwatts means a stronger signal strength which indicates a closer physical position to the AP since signal strengths decrease exponentially with distance.<font color="red">Find this in the book and cite it.</font> The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage.
+The higher the decibel-milliwatts means a stronger signal strength which indicates a closer physical position to the AP since signal strength decreases exponentially with distance.<font color="red">Find this in the book and cite it.</font> The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage.
 What would this plot look like if we moved the fixed location point on the other side of the building at (X=33, Y=3)?
@@
 plotBoxplotSignalStrength(df)
+# figure 3 - Signal strength distribution for a point at the other end of the building
 ```
@@
 At this point we know signal strength is different for the two access points.
-We don't know if this will effect our prediction accuracy.
-Now, let us analyzed the two APs signal strength distribution over angles.
+We don't know if this will affect our prediction accuracy.
+Now, let us analyze the two AP's signal strength distribution over angles.
+
+[&uarr;TOC](#TOC)
 
@@
 plotDensitySignalStrength(df)
+# figure 4 - Signal strength distribution across angles
 ```
@@
 Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle.
-This would suggest the median of each distribution varies by angle and attributes to a difference in perceived strength.
+This would suggest the median of each distribution varies by angle which can be attributed to a difference in perceived strength.
 This means the angle of the device contributes to the APs strength of signal.
@@
 #### Average Signal Strength Distributions
-Plot the signal strengh distribution for all locations through the average signal strenth for all locations by angle. Summary statistics for all possible locations by angle was created.
+Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle was created.
 
@@
 plotStdDevSignalStrength(df)
+# figure 5 - Standard deviation of average signal strength
 ```
@@
 This above figure shows that there is considerable overlap between the two distributions of signal strength. This suggest that there may be value in keeping both APs in the k-NN model.
+
+[&uarr;TOC](#TOC)
 
@@
 
+[&uarr;TOC](#TOC)
+
 ### 1.2.4 <a id='knn'>K-Nearest Neighbors</a>
-k-NN is a non-parametric machine learning algorithm that is instance-based learning as opposed to generalize learning.
-Our first implementation of k-NN learns based on teh number of neighbors within signal strength distance to the point in question from the training points.[2]
+k-NN is a non-parametric machine learning algorithm that is instance-based learning as opposed to generalized learning.
+Our first implementation of k-NN learns based on the number of neighbors within signal strength distance to the point in question from the training points.[2](#Ref02)
 
@@
 macs
+# Table 3 - Unique Access Points
 ```
@@
 tabonlineXYA
+# Table 5 - Test data counts per location and angle
 ```
@@
 names(onlineSummary)
+# Table 6 - Reshaped test data set
 ```
@@
   * If one angle is desired, then we will include angles which match the rounded orientation of the new observation  
-  * If we want two angles, two multiples of 45 degrees will be chosen in order to encapsulate the nw observation's orientation  
+  * If we want two angles, two multiples of 45 degrees will be chosen in order to encapsulate the new observation's orientation  
   * For three angles, the nearest 45 degree increment is chosen and either side will be used.
 
-Now, we analyze or three scenarios for both APs included, rejectedAP left out, and then chosenAP left out.
+Now, we analyze our three scenarios for both APs included, rejectedAP left out, and then chosenAP left out.
 Three neighbors were used for three angles.
-The accuracy for each scenario as determined by the sum of squares difference between estimated XY and actual XY were outputted as well.
+The accuracy for each scenario as determined by the sum of squares difference between estimated X,Y and actual X,Y were outputted as well.
 Cross-validation was not used at this time.
@@
 }
+# Table 7 - Accuracies for the various scenarios
 ```
@@
 
-We will now perform cross-validation with both APs to determine the best k.
-Three angles will be used.
-Future work would include parameterization of the angles from 1 to 3 to determine to determine the best m.
+We next performed cross-validation with both APs to determine the best k. Three angles were used. Future work would include parameterization of the angles from 1 to 3 to determine the best m.
 
@@
 plotSSErrors(err, K)
+# figure 6 - estimated ideal 'k'
 ```
@@
 
+[&uarr;TOC](#TOC)
+
 # <a id='Part2'> Part 2: Alternative k-Nearest Neighbors</a>
 
-## 2.1: Examination of Power-based search
+## 2.1: Examination of Power-based search {#ExamPower}
 The case study document mentioned the curvature in the scatterplots of signal strength versus distance, and suggested using a log function to attempt to remove the curve.
 Since the signal strength data are reported in dBm, they are already the result of a log function: $dBm = 10*log_{10}(power/1 mW)$; therefore, we converted the reported signal strength values back to a power value and examined the effect that had on location prediction.
-Many of the functions used had to be modified to include the power conversion: $power = 10^{(signal/10)}$ and to use the new field in the calculations.
+Many of the functions used previously had to be modified to include the power conversion: $power = 10^{(signal/10)}$ and to use the new field in the calculations.
 Examination of surface plots of power show that the power value tends to isolate the AP more definitively than signal strength because $power \propto 1/distance^2$.
 
-![Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
-![Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)
+![Figure 7 - Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
+
+![Figure 8 - Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)
 
@@
 However, when we simply substituted power for signal strength in the existing functions, the predictions using power were worse than using signal strength.
-![Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
+
+![Figure 9 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
 
@@
 
-| k |Mean distance error|Book-reported error|
+| k |Power-based error|Signal-strength-based error|
 |:-:|:-----------------:|:-----------------:|
@@
 
+*Table 8 - Comparison of predicted location errors between power-based and signal-strength-based approaches*
+
 Even though the power-based method better isolated the individual APs, we believe that the search algorithm should be redesigned to use all APs, instead of just one or two, due to the power drop-off.
@@
 
-## 2.2: Weighted K-Nearest Neighbors
+[&uarr;TOC](#TOC)
+
+##2.2: Investigating Weights {#ExamWeight}
+
+The description of the weighting procedure brought up some questions about the meaningfulness of the approach.
+This led to a side investigation of how the nearest-neighbors method might be affected by weights.
+
+The investigation involved exporting the offine summary dataset and the online dataset to CSV files for use in Excel.
+Then we looked at the process manually - taking the first report in the online dataset that matched our reference AP and looking for nearest neighbors in the offline summary dataset.
+The target record stated the true position to be 0, 0.05 (lower left corner of the building).
+The reference AP is within line of sight of the target device at coordinates 1,14.
+
+First, we selected only those records in the offline summary that matched the AP, then further filtered to those records that matched orientation (135 degrees).
+Plotting the measured signal strength for that AP and orientation, we can see that there is a fairly extensive swath of positions in which the target signal strength can occur.
+
+![*Figure 10 - Signal strength from AP @ 1,14*](SignalStrength1.png)
+
+Using the Excel function SMALL, we can find the k smallest differences - these would be the k nearest neighbors.
+
+|Record#|Sig Str Diff|Pos X|Pos Y|
+|:-----:|:----------:|:---:|:---:|
+| 3767  | 0.0196078  | 4   | 7   |
+|  357  | 0.0202128  | 1   | 2   |
+|  117  | 0.3846154  | 2   | 0   |
+| 3817  | 0.9354839  | 3   | 7   |
+|  505  | 1.0510204  | 1   | 3   |
+
+*Table 9 - Nearest neighbors for the target's signal strength*
+
+Avg X = 2.2; Avg Y = 3.8; True X = 0.0; True Y = 0.05; Error = 4.3477
+
+We can see that the target's reported signal strength occurs throughout the corridor as shown in the signal strength plot above.  Averaging those values pulls the estimate away from the true area.
+
+If we consider weighting the neighbors based on physical distance (as was suggested in the case study), the points closer to the AP will be weighted higher than the points further away, which would (in this particular example) pull the estimated location further away from the true location and closer to the AP, increasing the error further.
+
+The case study also hints at using multiple APs to do the neighbor selection.
+If, for example, we chose the AP at the opposite corner of the building (the lower right corner), and examined the same target report as before, we can see an additional problem:
+
+![*Figure 11 - Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)
+
+Now if we look for just simple differences in signal strength, we can get points for either AP that could be spread across the entire building.
+
+|Record#|Sig Str Diff|Pos X|Pos Y|AP|
+|:-----:|:----------:|:---:|:---:|:-:|
+|  670  | 0.0625     | 4   | 3   |1, 14|
+|  2568 | 0.0919540  | 15  | 8   |33.5, 2.8|
+|  856  | 0.1034483  | 11  | 4   |33.5, 2.8|
+|  936  | 0.1263158  | 12  | 5   |33.5, 2.8|
+|  408  | 0.1612903  | 0   | 7   |1, 14|
+
+*Table 10 - Nearest neighbors for the target's signal strength across two APs*
+
+Avg X = 8.4; Avg Y = 5.4; True X = 0.0; True Y = 0.05; Error = 9.959
+
+Again, weighting the neighbors on physical distance reinforces a location closer to the AP at 33.5, 2.8, which increases the error from the true location.
+
+One possible approach is to use the simultaneous measurements as a group and find nearest neighbors that exhibit similar patterns.
+For example the test record we have used here is part of a series of measurements taken at the same time and orientation.
+The highest average signal strength was against AP 97:8A (-43.25 dBm) and the lowest average signal strength was against AP 97:8D (-63 dBm).
+If we search for that pattern in the training data, we get a more reasonable set of neighbors:
+
+|Timestamp|97:8A signal|97:8D signal|Signal "Dist"|Pos X|Pos Y|
+|:-------:|:----------:|:----------:|:-----------:|:---:|:---:|
+|02:45:43 | -43 | -63 | 0.25  |1 | 5 |
+|02:14:34 | -44 | -63 | 0.75  |1 | 4 |
+|01:06:28 | -44 | -61 | 2.136 |0 | 1 |
+|06:31:15 | -41 | -63 | 2.25  |2 | 12|
+|00:35:27 | -41 | -64 | 2.462 |0 | 0 |
+
+*Table 11 - Nearest neighbors for the target's strongest and weakest signal strength*
+
+Avg X = 0.8; Avg Y = 4.4; True X = 0.0; True Y = 0.05; Error = 4.423
+
+Future work into this approach with appropriate weighting could prove to reduce location prediction errors.
+
+[&uarr;TOC](#TOC)
+
+## 2.3: Weighted K-Nearest Neighbors {#WtKNN}
+
+Prediction based on weighted X and Y &darr;
 
 ```{r}
-# create weights for neighbors based on signal strength and multiple times X and Y and sum neighbors
-# to get X and Y.
 estXY = predXY(newSignals = onlineSummary[ , 6:12],
-               newAngles = onlineSummary[ , 4],
-               offlineSummary, numAngles = 3, k = 5, weighted=TRUE)
-actXY = onlineSummary[ , c("posX", "posY")]
+               newAngles = onlineSummary[, 4],
+               offlineSummary, numAngles = 3, k = 5, weighted = TRUE, exp = .5)
+actXY = onlineSummary[, c("posX","posY")]
 calcError(estXY, actXY)
@@
 
+&darr; this is equivalent to mean SS error result of 210
+
+```{r}
+estXY = predXY(newSignals = onlineSummary[ , 6:12],
+               newAngles = onlineSummary[, 4],
+               offlineSummary, numAngles = 3, k = 5, weighted = TRUE, exp = .001)
+actXY = onlineSummary[, c("posX","posY")]
+calcError(estXY, actXY)
+```
+
+[&uarr;TOC](#TOC)
+
 ***
 # <a id='References'>References</a>
-[1] Nolan, Deborah; Data Science in R; 2015; Predicting Location via Indoor Positioning Systems; Chapter 1. 
-[2] Scikit-learn k-NN; http://scikit-learn.org/stable/modules/neighbors.html#nearest-neighbor-algorithms.
+* <a href='REF01'></a> [1] Nolan, Deborah; Data Science in R; 2015; Predicting Location via Indoor Positioning Systems; Chapter 1.
+
+* <a href='REF02'></a> [2] Scikit-learn k-NN; http://scikit-learn.org/stable/modules/neighbors.html#nearest-neighbor-algorithms.
 

### 2017-06-20 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Minor Corrections. I will work on this throughout tomorrow.
Adding 2.3 material from Ben's update

@@
 Three neighbors were used for three angles.
-The accuracy for each scenario as determined by the sum of squares difference between estimated X,Y and actual X,Y were outputted as well.
+The accuracy for each scenario as determined by the sum of squares difference between estimated X,Y and actual X,Y were output as well.
 Cross-validation was not used at this time.
@@
 ## 2.3: Weighted K-Nearest Neighbors {#WtKNN}
+In this exercise, we investigated using an alternative k-nearest prediction method using the weighted 
+K-nearest neighbors approach.
+The calculation of the Euclidean distance error was found to be 1194, whereas the Mean SS error was found to be 210.
+The Euclidean distance weighted k-nearest neighbors
+method had the highest error, while the Mean SS weighted K-nearest neighbors method had the smaller
+error.
+
+| k |Euclidean distance error|Mean SS error|
+|:-:|:-----------------:|:-----------------:|
+| |1194|210|
 

### 2017-06-21 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Added figure and others. See desc.
 I added a KNN figure, fixed the tables, removed comments, corrected some spelling/grammar, and added a "Future Study" section which describes some of what Matt was talking about

@@
 ```
-
 ### Table of Contents: {#TOC}
@@
   <li>[2.3 Weighted kNN](#WtKNN)</li>
+<li>[Future Study](#Future)</li>
 </ul>
@@
 
-<font color = "purple"><b>>>>>Everything in this color to be removed at final review<<<<</b></font>
-
-<font color = "purple"><b>
-- Discuss the goal of the case study
-</b></font>
-
 The case study in chapter 1 of the book *Data Science in R* by Nolan and Lang presented an indoor posiitoning system (IPS) for WiFi devices on a floor of a building at the University of Mannheim.
@@
 
-<font color = "red"><b>
-- Need example training input line
+This is an example of the training input:
 
-t=1139643118358;id=00:02:2D:21:0F:33;pos=0.0,0.0,0.0;degree=0.0;00:14:bf:b1:97:8a=-38,2437000000,3;00:14:bf:b1:97:90=-56,2427000000,3;00:0f:a3:39:e1:c0=-53,2462000000,3;00:14:bf:b1:97:8d=-65,2442000000,3;00:14:bf:b1:97:81=-65,2422000000,3;00:14:bf:3b:c7:c6=-66,2432000000,3;00:0f:a3:39:dd:cd=-75,2412000000,3;00:0f:a3:39:e0:4b=-78,2462000000,3;00:0f:a3:39:e2:10=-87,2437000000,3;02:64:fb:68:52:e6=-88,2447000000,1;02:00:42:55:31:00=-84,2457000000,1
-
-- Need table of training data: name, description, type, example
+t=1139643118358;id=00:02:2D:21:0F:33; pos=0.0,0.0,0.0; degree=0.0; 00:14:bf:b1:97:8a=-38,2437000000,3; 00:14:bf:b1:97:90=-56,2427000000,3; 00:0f:a3:39:e1:c0=-53,2462000000,3; 00:14:bf:b1:97:8d=-65,2442000000,3; 00:14:bf:b1:97:81=-65,2422000000,3; 00:14:bf:3b:c7:c6=-66,2432000000,3; 00:0f:a3:39:dd:cd=-75,2412000000,3; 00:0f:a3:39:e0:4b=-78,2462000000,3; 00:0f:a3:39:e2:10=-87,2437000000,3; 02:64:fb:68:52:e6=-88,2447000000,1; 02:00:42:55:31:00=-84,2457000000,1
 
 |Data Element|Description|Type|Example|
-|:-|:-|:-|:-|
-|t |timestamp (milliseconds since midnight 1/1/1970)|long integer|t=1139643118358 --> 2/11/2006 07:35:28|
+|:-----:|:------------------------------------:|:-----------------------------:|:---------:|
+| t | timestamp (milliseconds since midnight 1/1/1970) |long integer|t=1139643118358 --> 2/11/2006 07:35:28|
 |id|target device MAC address|6-octet sequence in hexadecimal|id=00:02:2D:21:0F:33|
@@
 
+***
+
 |Data Element|Description|Type|Example|
-|-|-|-|-|
+|:--|:----------------------------------------:|:--|:--:|
 |signal|The signal strength reported for the Access Point in dBm|integer|-38|
@@
 
-</b></font>
-
-<font color = "purple"><b>
-- Discuss the different Access points and MAC Addresses
-</b></font>
+***
 
@@
 The authors chose AP "00:0f:a3:39:e1:c0" (chosenAP) instead.
-This case study will evalute the author's decision and decide if the IPS system would be more accurate with the chosenAP, rejectedAP, or both APs.
-
-<font color = "purple"><b>
-- BRIEF overview of K-Nearest Neighbors
-</b></font>
+This case study will evalute the author's decision and decide if the IPS would be more accurate with the chosenAP, rejectedAP, or both APs.
 
@@
 
-<font color = "red"><b>
-- can we still a graphic from scikit-learn of k-NN model?
-</b></font>
+![Figure 1 - An example of k-Nearest Neighbors from scikit-learn [2](#REF02)](knnExample.png)
 
 In part 1 of the case study, the author's original decision to reject an AP is revisited to see if this was the best decision.
-In part 2, the prediction of X,Y based on the average of its neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions.
+In part 2, the prediction of X,Y based on the average of it's neighbors is revisited and modified to a weighted sum of X,Y to determine if this change will produce better predictions.
 Alternate weights will also be analyzed.
@@
 plotSignalMaps(c(chosenAP, rejectedAP))
-#  figure 1 - number of reports at each location for the rejected and chosen APs.
+#  figure 2 - number of reports at each location for the rejected and chosen APs.
 ```
 
-The maps show there were near equal amounts of training points captured for both APs. A visual comparison of 
-these two diagrams does not give us sufficient insight into which MAC address should be kept and which should be rejected. Now, let us examine the distribution strength of the APs.
+The maps show there were near equal amounts of training points captured for both APs.
+A visual comparison of these two diagrams does not give us sufficient insight into which MAC address should be kept and which should be rejected.
+Now, let us examine the distribution strength of the APs.
 
@@
 plotBoxplotSignalStrength(df)
-# figure 2 - Signal strength distribution across both Access Points
+# figure 3 - Signal strength distribution across both Access Points
 ```
 
-The higher the decibel-milliwatts means a stronger signal strength which indicates a closer physical position to the AP since signal strength decreases exponentially with distance.<font color="red">Find this in the book and cite it.</font> The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage.
+The higher the decibel-milliwatts (dBm) means a stronger signal strength which indicates a closer physical position to the AP since signal strength decreases exponentially with distance.
+The distribution shows AP "e1:c0" has a stronger signal strength at this location than "dd:cd." But, the two of them together cover a wider range of signal coverage.
 What would this plot look like if we moved the fixed location point on the other side of the building at (X=33, Y=3)?
@@
 plotBoxplotSignalStrength(df)
-# figure 3 - Signal strength distribution for a point at the other end of the building
+# figure 4 - Signal strength distribution for a point at the other end of the building
 ```
@@
 plotDensitySignalStrength(df)
-# figure 4 - Signal strength distribution across angles
+# figure 5 - Signal strength distribution across angles
 ```
-The above plot, shows the rejectedAP "dd:cd" has a higher signal density at lower strength while the acceptedAP "e1:c0" has a higher density coverage at a higher strength.
-The distributions show that many of the plots are normally distributed while some of them; i.e., "dd:cd" at 180 and "e1:c0" at 315 are heavily skewed or even bimodal.
+The above plot shows the rejectedAP "dd:cd" has a higher signal density at lower strength while the acceptedAP "e1:c0" has a higher density coverage at a higher strength.
+The distributions show many of the plots are normally distributed while some of them, (e.g., "dd:cd" at 180 and "e1:c0" at 315) are heavily skewed or even bimodal.
 Looking at each AP from top to bottom, there seems to exist an obvious pattern with respect to angle.
@@
 #### Average Signal Strength Distributions
-Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle was created.
+Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle were created.
 
@@
 plotStdDevSignalStrength(df)
-# figure 5 - Standard deviation of average signal strength
+# figure 6 - Standard deviation of average signal strength
 ```
 
-This above figure shows that there is considerable overlap between the two distributions of signal strength. This suggest that there may be value in keeping both APs in the k-NN model.
+This above figure shows there is considerable overlap between the two distributions of signal strength. This suggest there may be value in keeping both APs in the k-NN model.
 
@@
 
-Reshape the test data where each AP is in a column. 
+Let's Reshape the test data where each AP is in a column. 
 ```{r}
@@
 The author's choice to keep chosenAP was indeed better than rejectedAP, but keeping both of them is even better.
-
-<font color = "Red">talk about cross-validation</font>
 
@@
 plotSSErrors(err, K)
-# figure 6 - estimated ideal 'k'
+# figure 7 - estimated ideal 'k'
 ```
-Given the plot above, the best k shows 9, but this changes if you rerun the cross-validation. Experimentation with k yielded 5 with the lowest error of 210 so it was used below.
+Given the plot above, the best k shows 9, but this changes if you rerun the cross-validation. Experimentation with k yielded 5 with the lowest error of 210. Thus, it was used below.
 
@@
 
-![Figure 7 - Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
+![Figure 8 - Contour plot of power for measurements against AP '97:8D' at 0 degrees orientation](contour1.png)
 
-![Figure 8 - Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)
+![Figure 9 - Contour plot of power for measurements against AP 'E1:C0' at 0 degrees orientation](contour2.png)
 
@@
 
-![Figure 9 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
+![Figure 10 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
 
@@
 
-![*Figure 10 - Signal strength from AP @ 1,14*](SignalStrength1.png)
+![*Figure 11 - Signal strength from AP @ 1,14*](SignalStrength1.png)
 
@@
 
-![*Figure 11 - Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)
+![*Figure 12 - Signal strength from AP @ 33.5,2.8*](SignalStrength2.png)
 
@@
 The calculation of the Euclidean distance error was found to be 1194, whereas the Mean SS error was found to be 210.
-The Euclidean distance weighted k-nearest neighbors
-method had the highest error, while the Mean SS weighted K-nearest neighbors method had the smaller
-error.
+The Euclidean distance weighted k-nearest neighbors method had the highest error, while the Mean SS weighted K-nearest neighbors method had the smaller error.
 
@@
 [&uarr;TOC](#TOC)
+
+## <a id="Future">Future Study</a>
+To take our work further, something to consider for a future endeavour would be to analyze the results of the K-Nearest Neighbors analysis for each Chosen and Rejected Access Point by the use of an F-test.
+The results from this study would help to support our findings in this case study by accepting or rejecting the null hypothesis ($H_O$) of no statistical difference between the results.
+Once these results are discovered, an analysis of variance could be conducted as well to put the nail in the coffin.
 

### 2017-06-21 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
Cropped images to reduce white-space, minor wordsmithing, fixed links to reference documents

@@
 The IPS was trained and tested with data collected at the University with a hand-held device on a grid of 166 points spaced one meter apart in the hallways.
-The algorithm for the IPS is the K-Nearest Neighbors (k-NN) [1](#REF01).
+The algorithm for the IPS is the K-Nearest Neighbors (k-NN) [[1]](#REF01).
 
@@
 
+`Table 1 - Data Element Description for the training data`
+
 ***
@@
 |mode|The mode of operation for the Access Point (1 = Ad Hoc, 3 = Access Point)|integer|3|
+
+`Table 2 - Data Element Description for the Access Point measurements`
 
@@
 
-![Figure 1 - An example of k-Nearest Neighbors from scikit-learn [2](#REF02)](knnExample.png)
+![Figure 1 - An example of k-Nearest Neighbors from scikit-learn [[2]](#REF02)](knnExample.png)
 
@@
 
-The majority of the code in this case study was leveraged from [1](#Ref01). The code was refactored for readibility and modified for part 1 and 2 analysis.
+The majority of the code in this case study was leveraged from [[1]](#REF01). The code was refactored for readibility and modified for part 1 and 2 analysis.
 
@@
 
-```{r}
+```{r, fig.height=8}
 plotSignalMaps(c(chosenAP, rejectedAP))
-#  figure 2 - number of reports at each location for the rejected and chosen APs.
+#  Figure 2 - Number of reports at each location for the rejected and chosen APs.
 ```
@@
 plotBoxplotSignalStrength(df)
-# figure 3 - Signal strength distribution across both Access Points
+# Figure 3 - Signal strength distribution across both Access Points
 ```
@@
 plotBoxplotSignalStrength(df)
-# figure 4 - Signal strength distribution for a point at the other end of the building
+# Figure 4 - Signal strength distribution for a point at the other end of the building
 ```
@@
 plotDensitySignalStrength(df)
-# figure 5 - Signal strength distribution across angles
+# Figure 5 - Signal strength distribution across angles
 ```
@@
 plotStdDevSignalStrength(df)
-# figure 6 - Standard deviation of average signal strength
+# Figure 6 - Standard deviation of average signal strength
 ```
@@
 k-NN is a non-parametric machine learning algorithm that is instance-based learning as opposed to generalized learning.
-Our first implementation of k-NN learns based on the number of neighbors within signal strength distance to the point in question from the training points.[2](#Ref02)
+Our first implementation of k-NN learns based on the number of neighbors within signal strength distance to the point in question from the training points.[[2]](#REF02)
 
@@
 tabonlineXYA
-# Table 5 - Test data counts per location and angle
+# Table 4 - Test data counts per location and angle
 ```
@@
 names(onlineSummary)
-# Table 6 - Reshaped test data set
+# Table 5 - Reshaped test data set
 ```
@@
 }
-# Table 7 - Accuracies for the various scenarios
+# Table 6 - Accuracies for the various scenarios
 ```
@@
 plotSSErrors(err, K)
-# figure 7 - estimated ideal 'k'
+# Figure 7 - Estimated ideal 'k'
 ```
@@
 
-![Figure 10 - Red lines indicate distance error between predicted location (*) and true location (black squares) using k=4](errors.png)
+![Figure 10 - Red lines indicate distance error between predicted location (*) and true location (black dots) using k=4](errors.png)
 
@@
 
-*Table 8 - Comparison of predicted location errors between power-based and signal-strength-based approaches*
+*Table 7 - Comparison of predicted location errors between power-based and signal-strength-based approaches*
 
@@
 
-*Table 9 - Nearest neighbors for the target's signal strength*
+*Table 8 - Nearest neighbors for the target's signal strength*
 
@@
 
-*Table 10 - Nearest neighbors for the target's signal strength across two APs*
+*Table 9 - Nearest neighbors for the target's signal strength across two APs*
 
@@
 
-*Table 11 - Nearest neighbors for the target's strongest and weakest signal strength*
+*Table 10 - Nearest neighbors for the target's strongest and weakest signal strength*
 
@@
 ## 2.3: Weighted K-Nearest Neighbors {#WtKNN}
-In this exercise, we investigated using an alternative k-nearest prediction method using the weighted 
+In this exercise, we investigated using an alternative k-nearest neighbors prediction method using the weighted 
 K-nearest neighbors approach.
@@
 |:-:|:-----------------:|:-----------------:|
-| |1194|210|
+|5 |1194|210|
 
@@
 # <a id='References'>References</a>
-* <a href='REF01'></a> [1] Nolan, Deborah; Data Science in R; 2015; Predicting Location via Indoor Positioning Systems; Chapter 1.
+* <a id='REF01'></a> [1] Nolan, Deborah; Data Science in R; 2015; Predicting Location via Indoor Positioning Systems; Chapter 1.
 
-* <a href='REF02'></a> [2] Scikit-learn k-NN; http://scikit-learn.org/stable/modules/neighbors.html#nearest-neighbor-algorithms.
+* <a id='REF02'></a> [2] Scikit-learn k-NN; http://scikit-learn.org/stable/modules/neighbors.html#nearest-neighbor-algorithms.
+
+[&uarr;TOC](#TOC)
 
@@
 ```
+
+[&uarr;TOC](#TOC)

### 2017-06-21 MSDS7333_CaseStudy06_MBaldree_BBrock_TElkins_AKelly.Rmd
final review

@@
 A more thorough analysis would examine the distributions at all locations points, but time did not permit this in-depth analysis.
+Instead, we will later create a summary statistic to analyze.
 
@@
 
+Next, we plot the signal strength distributions for each AP per angle for better understanding at location (X=33, Y=3).
+
 ```{r}
@@
 ```
-This plot is similar to the previous distribution but with no outliers.
+This plot is similar to the previous distribution but with one outlier.
 At this point we know signal strength is different for the two access points.
-We don't know if this will affect our prediction accuracy.
+We do not know if this will affect our prediction accuracy.
 Now, let us analyze the two AP's signal strength distribution over angles.
@@
 #### Average Signal Strength Distributions
-Plot the signal strength distribution for all locations through the average signal strength for all locations by angle. Summary statistics for all possible locations by angle were created.
+Plot the signal strength distribution for all locations through the average signal strength for all locations by angle by creating a summary statistic for all possible locations.
 
@@
 ```
-The above output shows that each angle has test data meaning we can test all angle for accuracy.
+The above output shows that each angle has test data meaning we can test all angles for accuracy.
 
@@
 The above results show that including both APs yielded a better prediction accuracy followed by keeping chosenAP and then rejectedAP.
-The author's choice to keep chosenAP was indeed better than rejectedAP, but keeping both of them is even better.
+But was the difference significant or just do to chance?
+To answer this question, a null hypothesis should be performed.
 
-We next performed cross-validation with both APs to determine the best k. Three angles were used. Future work would include parameterization of the angles from 1 to 3 to determine the best m.
+Our null hypothesis is that all scenarios yield the same error. Our alternative hypothesis is that at least one scenario yields a different error from the others.
+
+* Ho: None Error = ChosenAP Error = RejectedAP Error
+* Ha: At least one scenario is different.
+
+The next step is test our hypothesis with an ANOVA test. This step is left for future work. Therefore, we cannot reject Ho at this time.
+
+We will proceed with our analysis with both APs and perform a cross-validation to determine the best k.
+Three angles were used.
+Future work would include parameterization of the angles from 1 to 3 to determine the best m.
 
@@
 
-<font color = "Red">provide some feel good about part 1.
-we concluded that keeping both APs yields better accuracy with a k-NN of 5 with 3 angles.
-We did not iterate of the angles to determine best parameter (arg!).
-The training data neighbors to the test data was determined by the signal strength distance and then the neighbors average X,Y was used to predict the test point location.
-</font>
+Our analysis of three scenarios that included both ChosenAP and RejectedAP and each exclused showed that there might be a potential gain from including both APs.
+Indeed the accuracy results showed including both APs yielded lower errors.
+However, this difference in errors was not analyzed for significance and was left for future work.
+A cross-validation for both APs scenario was performed to find optimal k.
+Analysis for optimal angle, m, was not peformed and left for future work.
 

### 2017-06-22 MSDS7333-baldree-case6-book.R
Prepped for Final Submission

@@ -1,705 +1,0 @@
-# MSDS7333 - Quantifying the World
-# Matt Baldree
-# Tom Elkins
-# Ben Brock
-# Austin Kelly
-#
-# Code from Data Science in R, Case Study 1
-
-#### 1.2 Raw Data ####
-options(digits = 2)
-# readlines for examination
-txt = readLines("offline.final.trace.txt")
-# how many lines begin with #? 5312
-sum(substr(txt, 1, 1) == "#")
-# how many lines in the file? 151,392
-length(txt)
-# split element on ';' then ='
-unlist(lapply(strsplit(txt[4],";")[[1]],function(x) sapply(strsplit(x,"=")[[1]],strsplit,",")))
-# prereview tokens from 4th line in file split on ';'
-strsplit(txt[4], ";")[[1]]
-# create tokens of line 4
-tokens = strsplit(txt[4], "[;=,]")[[1]]
-# view the first 10 tokens of line 4
-tokens[1:10]
-# grab the values
-tokens[c(2, 4, 6:8, 10)]
-# recorded signals within an observation
-tokens[-(1:10)]
-# unravel signals and turn into a 11x10 matrix
-tmp = matrix(tokens[-(1:10)], ncol = 4, byrow = TRUE)
-mat = cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp), ncol = 6, byrow = TRUE), tmp)
-dim(mat)
-
-# create function to process a line in the file
-processLine = function(x)
-{
-  tokens = strsplit(x, "[;=,]")[[1]]
-  tmp = matrix(tokens[-(1:10)], ncol = 4, byrow = TRUE)
-  cbind(matrix(tokens[c(2, 4, 6:8, 10)], nrow = nrow(tmp), ncol = 6, byrow = TRUE), tmp)
-}
-# apply function to several lines
-tmp = lapply(txt[4:20], processLine)
-# simplify
-sapply(tmp, nrow)
-# stack matrices together
-offline = as.data.frame(do.call("rbind", tmp))
-dim(offline)
-
-# grab all lines that do not begin with '#'
-lines = txt[ substr(txt, 1, 1) != "#" ]
-# process them; warning messages will appear in output that data exceed size of matrix
-tmp = lapply(lines, processLine)
-
-# modify function
-processLine = function(x)
-{
[…]

### 2017-06-22 MSDS7333-baldree-case6.r

@@ -1,223 +1,0 @@
-# MSDS7333 - Quantifying the World
-# Matt Baldree
-# Tom Elkins
-# Ben Brock
-# Austin Kelly
-#
-# Assignment:
-# 1. Conduct a more thorough data analysis into these two MAC addresses including determining locations
-# by using data corresponding to both MAC addresses.
-#   Which of these two MAC addresses should be used and which should not be used for RTLS?
-#   Which MAC address yields the best prediction of location?
-#   Does using data for both MAC addresses simultaneously yield more, or less, accurate prediction of location?
-#
-# 2. Implement alternative k-nearest prediction method using weights on received signal strength.
-#   For what range of values of weights are you able to obtain better prediction values than for the unweighted approach?
-#     Use calcError() to compare approaches.
-#
-# Write up assignment in an Ipython Notebook with an introduction, explanation of approaches and outputs.
-#
-
-# two digit precision for printing and formatting
-options(digits = 2)
-options(error = recover, warn = 1)
-# include necessary functions
-source("MSDS7333-baldree-case6-fx.r", print.eval=TRUE)
-
-#### Part 1: Access point MAC address analysis of 00:0f:a3:39:e1:c0 and 00:0f:a3:39:dd:cd ####
-
-# read offline data
-offline = readData()
-
-# create factor for all unique xy combinations
-offline$posXY = paste(offline$posX, offline$posY, sep = "-")
-
-# access points (AP) addresses in question
-# note that both devices are Alpha Networks devices because they begin with '00:0f'
-chosenAP = "00:0f:a3:39:e1:c0"
-rejectedAP = "00:0f:a3:39:dd:cd"
-remainingAP = unique(offline[!offline$mac %in% c(chosenAP, rejectedAP),]$mac)
-
-#### Signal Collection Map ####
-
-# plot the locations and count of signals recorded for both access points
-# we would plot the APs on the map but we do not know their X,Y locations
-# result: map looks equivalent
-plotSignalMaps(c(chosenAP, rejectedAP))
-
-#### Signal Strength Distribution ####
-
-# plot signal strength distribution for each angle per AP
-# we are using the same stationary point of 2, 12 as the book
-# result: rejected AP has a lower signal strength for the X,Y than chosen AP.
-df = subset(offline, posX == 2 & posY == 12 & !(mac %in% remainingAP), c(signal, angle, mac))
-plotBoxplotSignalStrength(df)
-
[…]
